---
title: "Polar Star Quantitative Commodity Fund"
author: "Dr Mauritz van den Worm, PhD"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  revealjs::revealjs_presentation:
    reveal_options:
      minScale: 1.0
      maxScale: 1.0
      slideNumber: true
    includes:
      in_header: addlogo.html
    theme: simple
    highlight: pygments
    center: true
    mathjax: "https://cdn.mathjax.org/mathjax/latest/MathJax.js?config=TeX-AMS-MML_HTMLorMML"
    self_contained: false
    reveal_plugins: ["menu"]
    css: custom.css
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)

library(data.table)
library(ggplot2)
library(plotly)
library(rhandsontable)
library(knitr)
library(readr)
library(purrr)
library(wavethresh)
library(extrafont)
library(scales)
library(PortfolioAnalytics)
library(formattable)
library(tseries)

loadfonts(device = "win")


```





# Manual vs Automated Trading  {data-background="img/background_mtns_1.png"}


##  {data-background="img/BlueBackgroundWithOrange.png"}

*How has the increase of Automated Trading Systems (ATS) influenced the futures market?*

- CME transaction data identitifies ATS with the 1028 tag
- Data available from November 2012 to present
- Consider data form 2012 to 2016
- Use gradient plots to highlight the changes
- Graphs are interactive






## MAN vs ATS - Group  {data-background="img/BlueBackgroundWithOrange.png"}


```{r}

manvsautbygroup <- fread("./data/manualvsautomatedbyproductgroup.csv")
manvsautbygroup <- manvsautbygroup[, c(1,3:6), with = FALSE]
plotdata <- melt(manvsautbygroup, id.vars = c("ProductGroupName", "Period"))

pic <- ggplot(plotdata, aes(x=factor(Period), y=value)) +
  geom_line(aes(group=ProductGroupName, color=ProductGroupName)) +
  geom_point(aes(color=ProductGroupName)) +
  facet_wrap(~variable) +
  ylab("Percentage of trades") +
  xlab("") +
  theme(
    plot.background = element_rect(fill = rgb(242,242,242, maxColorValue = 255)),
    panel.background = element_rect(fill = "white"),
    panel.grid.major.y = element_line(size = 0.5, linetype = 'solid',colour = "gray"),
    panel.grid.major.x = element_line(size = 0.5, linetype = 'dashed',colour = "gray"),
    axis.title.x = element_text(colour = rgb(5,61,96, maxColorValue = 255), size = 16),
    axis.title.y = element_text(colour = rgb(5,61,96, maxColorValue = 255), size = 16),
    plot.title = element_text(colour = rgb(5,61,96, maxColorValue = 255), size = 18, face = "bold"),
    axis.text = element_text(size=16),
    legend.key = element_rect(fill = rgb(242,242,242, maxColorValue = 255)),
    legend.background=element_blank(),
    legend.title=element_blank(),
    text=element_text(family="Century Gothic")
  )


ggplotly(pic)


```




## MAN vs ATS - Agriculture  {data-background="img/BlueBackgroundWithOrange.png"}


```{r}

agriculture <- fread("./data/manualvsautomatedbyproductsubgroup.csv")
agriculture <- agriculture[ProductGroup == "Agriculture"]
plotdata <- melt(agriculture, id.vars = c("ProductGroup", "Subgroup", "Period"))

pic <- ggplot(plotdata, aes(x=factor(Period), y=value)) +
  geom_line(aes(group=Subgroup, color=Subgroup)) +
  geom_point(aes(color=Subgroup)) +
  facet_wrap(~variable) +
  ylab("Percentage of trades") +
  xlab("") +
  theme(
    plot.background = element_rect(fill = rgb(242,242,242, maxColorValue = 255)),
    panel.background = element_rect(fill = "white"),
    panel.grid.major.y = element_line(size = 0.5, linetype = 'solid',colour = "gray"),
    panel.grid.major.x = element_line(size = 0.5, linetype = 'dashed',colour = "gray"),
    axis.title.x = element_text(colour = rgb(5,61,96, maxColorValue = 255), size = 16),
    axis.title.y = element_text(colour = rgb(5,61,96, maxColorValue = 255), size = 16),
    plot.title = element_text(colour = rgb(5,61,96, maxColorValue = 255), size = 18, face = "bold"),
    axis.text = element_text(size=16),
    legend.key = element_rect(fill = rgb(242,242,242, maxColorValue = 255)),
    legend.background=element_blank(),
    legend.title=element_blank(),
    text=element_text(family="Century Gothic")
  )

ggplotly(pic)

```




## MAN vs ATS - Energy  {data-background="img/BlueBackgroundWithOrange.png"}


```{r}

energy <- fread("./data/manualvsautomatedbyproductsubgroup.csv")
energy <- energy[ProductGroup == "Energy"]
plotdata <- melt(energy, id.vars = c("ProductGroup", "Subgroup", "Period"))

pic <- ggplot(plotdata, aes(x=factor(Period), y=value)) +
  geom_line(aes(group=Subgroup, color=Subgroup)) +
  geom_point(aes(color=Subgroup)) +
  facet_wrap(~variable) +
  ylab("Percentage of trades") +
  xlab("") +
  theme(
    plot.background = element_rect(fill = rgb(242,242,242, maxColorValue = 255)),
    panel.background = element_rect(fill = "white"),
    panel.grid.major.y = element_line(size = 0.5, linetype = 'solid',colour = "gray"),
    panel.grid.major.x = element_line(size = 0.5, linetype = 'dashed',colour = "gray"),
    axis.title.x = element_text(colour = rgb(5,61,96, maxColorValue = 255), size = 16),
    axis.title.y = element_text(colour = rgb(5,61,96, maxColorValue = 255), size = 16),
    plot.title = element_text(colour = rgb(5,61,96, maxColorValue = 255), size = 18, face = "bold"),
    axis.text = element_text(size=16),
    legend.key = element_rect(fill = rgb(242,242,242, maxColorValue = 255)),
    legend.background=element_blank(),
    legend.title=element_blank(),
    text=element_text(family="Century Gothic")
  )

ggplotly(pic)

```





## Trade Participation  {data-background="img/BlueBackgroundWithOrange.png"}


```{r}

mostactive <- fread("./data/manualvsautomatedmostactive.csv")
mostactive <- mostactive[ProductGroup %in% c("Agriculture", "Energy", "Metals")]
plotdata <- melt(mostactive, id.vars = c("ProductGroup", "Subgroup", "Futures Product", "Period"))

pic <- ggplot(plotdata, aes(x=factor(Period), y=value)) +
  geom_line(aes(group=`Futures Product`, color=`Futures Product`)) +
  geom_point(aes(color=`Futures Product`)) +
  facet_wrap(~variable) +
  ylab("Percentage of trades") +
  xlab("") +
  theme(
    plot.background = element_rect(fill = rgb(242,242,242, maxColorValue = 255)),
    panel.background = element_rect(fill = "white"),
    panel.grid.major.y = element_line(size = 0.5, linetype = 'solid',colour = "gray"),
    panel.grid.major.x = element_line(size = 0.5, linetype = 'dashed',colour = "gray"),
    axis.title.x = element_text(colour = rgb(5,61,96, maxColorValue = 255), size = 16),
    axis.title.y = element_text(colour = rgb(5,61,96, maxColorValue = 255), size = 16),
    plot.title = element_text(colour = rgb(5,61,96, maxColorValue = 255), size = 18, face = "bold"),
    axis.text = element_text(size=16),
    legend.key = element_rect(fill = rgb(242,242,242, maxColorValue = 255)),
    legend.background=element_blank(),
    legend.title=element_blank(),
    text=element_text(family="Century Gothic")
  )

ggplotly(pic)

```


## Where are the opportunities?  {data-background="img/BlueBackgroundWithOrange.png"}

- Limited, Local, Spectrum
    - Less liquid parts of the curve
    - Changes to macro
    - Concentrated risk on asymmetric risk/reward trades
    
</br>

- Quantitative
    - Harvest systematic alpha from large universe of commodities
    - Rule based decision making process
    - Strategies inspired by discretional methodology


## Literature  {data-background="img/BlueBackgroundWithOrange.png"}

- <a href = "https://www.cmegroup.com/education/files/Algo_and_HFT_Trading_0610.pdf" target="_blank">Algorithmic trading and Market Dynamics</a>

- <a href = "http://www.cftc.gov/idc/groups/public/@economicanalysis/documents/file/oce_automatedtrading.pdf" target="_blank">Automated Trading in Futures Markets, Richard Haynes and John S. Roberts</a>

- <a href = "http://www.cftc.gov/idc/groups/public/@economicanalysis/documents/file/oce_automatedtrading_update.pdf" target="_blank">Automated Trading in Futures Markets - Update, Richard Haynes and John S. Roberts</a>







# Grinold's Fundamental Law of Active Portfolio Management {data-background="img/background_mtns_1.png"}

## Law of active portfolio management {data-background="img/BlueBackgroundWithOrange.png"}


$$
\text{Performance} = \text{Skill} \times \sqrt{\text{Breadth}}
$$



## Suppose we are in a coin flipping casino {data-background="img/BlueBackgroundWithOrange.png"}

- Flip coins in stead of futures
- The coin is biased - $P(\text{heads}) = 0.51$



## How does the betting work? {data-background="img/BlueBackgroundWithOrange.png"}

- We have 1000 coins 
- The minimum wager is 1 coin
- If you win you gain 1 coin
- If you loose you loose 1 coin
- There are 1000 tables with coin wagers
- Games runs in parallel


## What is the optimal way to allocate coins? {data-background="img/BlueBackgroundWithOrange.png"}

Two extremes

- Bet 1000 coins on one coin flip
- Bet 1 coin on 1000 coin flips


## Expected Return {data-background="img/BlueBackgroundWithOrange.png"}

- Single bet:   $0.51 \times 1000 + 0.49 \times (-1000) = 20$

- Multi bet:   $1000 \times [0.51  + 0.49 \times (-1)]  = 20$

The same expected return


## Risk - Probability to lose it all: {data-background="img/BlueBackgroundWithOrange.png"}

- Single bet: 49%

- Multi bet: $0.49 \times 0.49 \times \dots \times 0.49 = 0.49^{1000} \approx 0$



## Risk - Standard Deviation: {data-background="img/BlueBackgroundWithOrange.png"}

- One coin per table

$$
\text{risk} := \text{std}\left\{1,-1,-1,1, \dots, 1 \right\} = 1
$$



- One 1000 coin bet, 999 zero coin bets


$$
\begin{align}
  \text{risk} &:= \text{std}\left\{1000,0,0,0, \dots, 0 \right\} = 31.62 \\
  \text{risk} &:= \text{std}\left\{-1000,0,0,0, \dots, 0 \right\} = 31.62
\end{align}
$$




## Coin Flip Casino - Reward/Risk {data-background="img/BlueBackgroundWithOrange.png"}

- Just like Sharpe Ratio

- Single bet:   $\text{SR}_{\text{single}} = \frac{20}{31.62} =0.63$

- Multi bet:   $\text{SR}_{\text{multiple}} = \frac{20}{1} =20$



## Coin flipping casino - Observation {data-background="img/BlueBackgroundWithOrange.png"}

- $20 = 0.63 \times \sqrt{1000}$
 
- $\text{SR}_{\text{multiple}} = \text{SR}_{\text{single}} \times \sqrt{\text{Bets}}$

- $\text{Performance} = \text{Skill} \times \sqrt{\text{Breadth}}$



## How does this apply to commodity futures? {data-background="img/BlueBackgroundWithOrange.png"}

- We use insights gained from years of fundamental trading to inspire bespoke quantitative strategies that are applied to a large collection of commodity markets

- We increase breadth or diversification by
    - how,
    - what and
    - when we trade


## When we trade {data-background="img/BlueBackgroundWithOrange.png"}

```{r}

tradeData <- as.data.table(read_csv("C:/Users/mauritz/Box Sync/Polar Star - Company Folder/1. Polar Star/Analysis/Analysis and Research/code/Quant Start Ideas/PS trend/tradeData.csv"))

tradeData[, monthcut := cut(month, breaks = c(0,3,6,9,12,15,18,21,24))]

```


```{r}

plotdata <- tradeData[fund != "Spectrum", .N, by = .(fund, monthcut)]
plotdata[, total := sum(N), by = fund]
plotdata[, percentage := N/total * 100]
plotdata[, monthnumber := as.integer(monthcut) * 3]

ggplot(plotdata, aes(x=factor(monthnumber), y=percentage)) +
  geom_bar(stat = "identity", fill = rgb(5,61,96, maxColorValue = 255)) + 
  xlab("Month Ranges") +
  ylab("Percentage of Trades") +
  geom_text(aes(label = round(percentage,1)), nudge_y = 1, colour = rgb(5,61,96, maxColorValue = 255)) +
  facet_wrap(~fund) +
  ggtitle("Percentage of trades with months until expiry") +
  theme(
    plot.background = element_rect(fill = rgb(242,242,242, maxColorValue = 255)),
    panel.background = element_rect(fill = "white"),
    panel.grid.major.y = element_line(size = 0.5, linetype = 'solid',colour = "gray"),
    panel.grid.major.x = element_line(size = 0.5, linetype = 'dashed',colour = "gray"),
    axis.title.x = element_text(colour = rgb(5,61,96, maxColorValue = 255), size = 16),
    axis.title.y = element_text(colour = rgb(5,61,96, maxColorValue = 255), size = 16),
    plot.title = element_text(colour = rgb(5,61,96, maxColorValue = 255), size = 18, face = "bold"),
    axis.text = element_text(size=12),
    text=element_text(family="Century Gothic"),
    strip.text.x = element_text(size = 14, colour = rgb(5,61,96, maxColorValue = 255))
  )


```


## What we trade {data-background="img/BlueBackgroundWithOrange.png"}

```{r}

tradeData[, comdty := tstrsplit(contractID, "_", keep = 1)]
tradeData[, month := month(`Trade Date`)]
tradeData[, year := year(`Trade Date`)]

plotdata <- tradeData[fund != "Spectrum", .(comdty, fund, year, month)]
setkey(plotdata, NULL)
plotdata <- unique(plotdata)


plotdata <- plotdata[, .N, by = .(fund, year, month)]
plotdata[, date := paste(year, month, 1, sep = "-")]
plotdata[, date := as.Date(date)]

ggplot(plotdata[date >= "2018-07-01"], aes(x=date, y= N)) +
  geom_bar(stat = "identity", fill = rgb(5,61,96, maxColorValue = 255)) +
  facet_wrap(~fund) +
  scale_y_continuous(breaks = pretty_breaks(n = 10)) +
  geom_text(aes(label = N), nudge_y = 1) + 
  ggtitle("Unique commodities traded in each fund since July") +
  ylab("Number of unique commodities") +
  xlab("Month") +
  theme(
    plot.background = element_rect(fill = rgb(242,242,242, maxColorValue = 255)),
    panel.background = element_rect(fill = "white"),
    panel.grid.major.y = element_line(size = 0.5, linetype = 'solid',colour = "gray"),
    panel.grid.major.x = element_line(size = 0.5, linetype = 'dashed',colour = "gray"),
    axis.title.x = element_text(colour = rgb(5,61,96, maxColorValue = 255), size = 16),
    axis.title.y = element_text(colour = rgb(5,61,96, maxColorValue = 255), size = 16),
    plot.title = element_text(colour = rgb(5,61,96, maxColorValue = 255), size = 16, face = "bold"),
    axis.text = element_text(size=12),
    text=element_text(family="Century Gothic"),
    strip.text.x = element_text(size = 14, colour = rgb(5,61,96, maxColorValue = 255))
  )



```


## Literature {data-background="img/BlueBackgroundWithOrange.png"}

- <a href = "http://jpm.iijournals.com/content/15/3/30" target="_blank">The fundamental law of active management - Richard C. Grinold</a>
- <a href = "https://www.jpmorgan.com/cm/BlobServer/Extending_the_Fundamental_Law_of_Investment_Management_.pdf?blobkey=id&blobwhere=1158630145176&blobheader=application%2Fpdf&blobheadername1=Cache-Control&blobheadervalue1=private&blobcol=urldata&blobtable=MungoBlobs" target="_blank">Extending the Fundamental Law
of Investment Management - JP Morgan US Equity Quantitative Research</a>



# Technical considerations when trading futures systematically  {data-background="img/background_mtns_1.png"}

## {data-background="img/BlueBackgroundWithOrange.png"}

- Continuous Futures Price Series
    - We require long time series data
    - Futures expire too soon to gather sufficient data
    - How do you handle rolls?

</br>    
    
- Non-stationarity of Price Data
    - Time series data can only reliably be forecasted if stationary
    - Machine Learning algorithms are designed for stationary features
    - How do we create stationary data?


## Continuous Futures Price Series {data-background="img/BlueBackgroundWithOrange.png"}


```{r}

contPrices <- as.data.table(read_csv("C:/Users/mauritz/Box Sync/Polar Star - Company Folder/1. Polar Star/Analysis/Analysis and Research/CalendarPortfolio/contCalendarFracDiff.csv"))

```


```{r}

pic <- ggplot(contPrices[comdty == "C" & tenure == "F3"], 
              aes(x=index, y=frontprice, text = paste('Price: ', frontprice,
                                         '<br>Date: ', as.Date(index)))) +
  geom_line(aes(group = frontID), colour = rgb(5,61,96, maxColorValue = 255)) +
  theme_bw() +
  xlab("") +
  ylab("Price") +
  ggtitle("Corn Continuous Price Curve with rolls indicated by discontinuities") +
  scale_x_date(breaks = pretty_breaks(n=12)) +
  theme(
    plot.background = element_rect(fill = rgb(242,242,242, maxColorValue = 255)),
    panel.background = element_rect(fill = "white"),
    panel.grid.major.y = element_line(size = 0.5, linetype = 'solid',colour = "gray"),
    panel.grid.major.x = element_line(size = 0.5, linetype = 'dashed',colour = "gray"),
    axis.title.x = element_text(colour = rgb(5,61,96, maxColorValue = 255), size = 16),
    axis.title.y = element_text(colour = rgb(5,61,96, maxColorValue = 255), size = 16),
    plot.title = element_text(colour = rgb(5,61,96, maxColorValue = 255), size = 14, face = "bold"),
    axis.text = element_text(size=14),
    text=element_text(family="Century Gothic"),
    legend.key = element_rect(fill = rgb(242,242,242, maxColorValue = 255)),
    legend.background=element_blank(),
    legend.title=element_blank(),
    legend.position="bottom"
  )


ggplotly(pic) %>%
  layout(legend = list(orientation = "h", x = 0.4, y = -0.2))

```


## Continuous Futures Curves {data-background="img/BlueBackgroundWithOrange.png"} 


```{r}

plotdata <- contPrices[comdty == "C" & tenure == "F3", .(index, frontID,  `Raw Price` = frontprice, `Roll Adjusted Price` = `adj frontprice`)]
plotdata <- melt(plotdata, id.vars = c("index", "frontID"))

pic <- ggplot(plotdata, aes(x=index, y=value, text = paste('Price: ', value,
                                         '<br>Date: ', as.Date(index)))) +
  geom_line(aes(group = variable, color = variable)) +
  theme_bw() +
  xlab("") +
  ylab("Price") +
  ggtitle("Corn Continuous Roll Adjusted Price Curve") +
  scale_x_date(breaks = pretty_breaks(n=10)) +
  scale_color_manual(values = c(rgb(5,61,96, maxColorValue = 255), rgb(255, 90, 52, maxColorValue = 255))) +
  theme(
    plot.background = element_rect(fill = rgb(242,242,242, maxColorValue = 255)),
    panel.background = element_rect(fill = "white"),
    panel.grid.major.y = element_line(size = 0.5, linetype = 'solid',colour = "gray"),
    panel.grid.major.x = element_line(size = 0.5, linetype = 'dashed',colour = "gray"),
    axis.title.x = element_text(colour = rgb(5,61,96, maxColorValue = 255), size = 16),
    axis.title.y = element_text(colour = rgb(5,61,96, maxColorValue = 255), size = 16),
    plot.title = element_text(colour = rgb(5,61,96, maxColorValue = 255), size = 14, face = "bold"),
    axis.text = element_text(size=16),
    text=element_text(family="Century Gothic"),
    legend.key = element_rect(fill = rgb(242,242,242, maxColorValue = 255)),
    legend.background=element_blank(),
    legend.title=element_blank(),
    legend.position="bottom"
  )


ggplotly(pic)%>%
  layout(legend = list(orientation = "h", x = 0.4, y = -0.2))

```


## Stationarity {data-background="img/BlueBackgroundWithOrange.png"} 

```{r}

plotdata <- contPrices[comdty == "C" & 
                         tenure == "F3" &
                         index >= "1990-01-01" &
                         index < "1992-01-01", .(`Non Stationary` = frontRP)]
plotdata[, index := 1:nrow(plotdata)]

plotdata[, Stationary := diff(`Non Stationary`)/`Non Stationary`]
plotdata[, Stationary := guyrot(Stationary, 1)]
plotdata[1, Stationary := NA]
plotdata[, Stationary := Stationary * 10]
plotdata[, `Non Stationary` := `Non Stationary` * 2 - 1]

plotdata <- melt(plotdata, id.vars = c("index"))

adfData <- plotdata[!is.na(value), adf.test(value), by = variable]
adfData <- adfData[,.(variable, statistic, p.value)]
adfData[, statistic := round(statistic, 3)]
adfData[, p.value := round(p.value, 3)]

ggplot(plotdata, aes(x=index, y=value)) +
  geom_line(aes(group = variable, color = variable)) +
  theme_bw() +
  xlab("") +
  ylab("Price") +
  geom_text(label = paste("adf:", adfData[variable == "Non Stationary", statistic]), x = 450, y = 1, color = rgb(5,61,96, maxColorValue = 255)) +
  geom_text(label = paste("adf:", adfData[variable == "Stationary", statistic]), x = 450, y = 0.25, color = rgb(255, 90, 52, maxColorValue = 255)) +
  ggtitle("Stationary and Non Stationary Time Series - Threshold = -2.8623") +
  #scale_x_date(breaks = pretty_breaks(n=10)) +
  scale_color_manual(values = c(rgb(5,61,96, maxColorValue = 255), rgb(255, 90, 52, maxColorValue = 255))) +
  theme(
    plot.background = element_rect(fill = rgb(242,242,242, maxColorValue = 255)),
    panel.background = element_rect(fill = "white"),
    panel.grid.major.y = element_line(size = 0.5, linetype = 'solid',colour = "gray"),
    panel.grid.major.x = element_line(size = 0.5, linetype = 'dashed',colour = "gray"),
    axis.title.x = element_text(colour = rgb(5,61,96, maxColorValue = 255), size = 16),
    axis.title.y = element_text(colour = rgb(5,61,96, maxColorValue = 255), size = 16),
    plot.title = element_text(colour = rgb(5,61,96, maxColorValue = 255), size = 14, face = "bold"),
    axis.text = element_text(size=16),
    text=element_text(family="Century Gothic"),
    legend.key = element_rect(fill = rgb(242,242,242, maxColorValue = 255)),
    legend.background=element_blank(),
    legend.title=element_blank(),
    legend.position="bottom"
  )



```

## Stationarity {data-background="img/BlueBackgroundWithOrange.png"} 

How to obtain a stationary time series

</br>

- Traditional
    - Price differences
    - Returns
    - Memory loss

</br>

- Modern
    - Fractional differences
    - Memory present

## Stationarity {data-background="img/BlueBackgroundWithOrange.png"} 


```{r}

fracdiffData <- as.data.table(read_csv("C:/Users/mauritz/Box Sync/Polar Star - Company Folder/1. Polar Star/Analysis/Analysis and Research/CalendarPortfolio/calendarFractionalDifferences.csv"))

setnames(fracdiffData, "X1", "d")

```


```{r}

plotdata <- fracdiffData[comdty == "C" & type == "front"]
plotdata[, adjcorr := corr * 60 - 59]

ggplot(plotdata, aes(x = d)) +
  geom_line(aes(y = adfStat), color = rgb(5,61,96, maxColorValue = 255), size = 1.2) +
  theme_bw() +
  xlab("Degree of Difference") +
  ylab("adf Statistic") +
  geom_hline(yintercept = -2.8623, color = rgb(255, 90, 52, maxColorValue = 255), size = 1.2) +
  geom_line(aes(y = adjcorr), linetype = "dashed", color = rgb(5,61,96, maxColorValue = 255), size = 1.2) +
  scale_y_continuous(sec.axis = sec_axis(~. /-60, name = "Correlation")) +
  ggtitle("Corn adf Statistic and Correlations with varying Degree of Difference") +
  #scale_color_manual(values = c(rgb(5,61,96, maxColorValue = 255), rgb(255, 90, 52, maxColorValue = 255))) +
  theme(
    plot.background = element_rect(fill = rgb(242,242,242, maxColorValue = 255)),
    panel.background = element_rect(fill = "white"),
    panel.grid.major.y = element_line(size = 0.5, linetype = 'solid',colour = "gray"),
    panel.grid.major.x = element_line(size = 0.5, linetype = 'dashed',colour = "gray"),
    axis.title.x = element_text(colour = rgb(5,61,96, maxColorValue = 255), size = 16),
    axis.title.y = element_text(colour = rgb(5,61,96, maxColorValue = 255), size = 16),
    plot.title = element_text(colour = rgb(5,61,96, maxColorValue = 255), size = 14, face = "bold"),
    axis.text = element_text(size=16),
    text=element_text(family="Century Gothic"),
    legend.key = element_rect(fill = rgb(242,242,242, maxColorValue = 255)),
    legend.background=element_blank(),
    legend.title=element_blank(),
    legend.position="bottom"
  )

```



## Stationarity {data-background="img/BlueBackgroundWithOrange.png"} 

```{r}


plotdata <- contPrices[comdty == "C" & tenure == "F3", .(index, frontID,  `ETF Price` = frontRP, `Fractional Difference` = frontFD)]

plotdata[, `Differenced Price` := diff(`ETF Price`)/`ETF Price`]
plotdata[, `Differenced Price` := guyrot(`Differenced Price`, 1)]
plotdata[1, `Differenced Price` := NA]

plotdata <- melt(plotdata, id.vars = c("index", "frontID"))


pic <- ggplot(plotdata, aes(x=index, y=value, text = paste('Price: ', value,
                                         '<br>Date: ', as.Date(index)))) +
  geom_line(aes(group = variable, color = variable)) +
  theme_bw() +
  xlab("") +
  ylab("Price") +
  ggtitle("Corn ETF and Fractionally Differenced Price Curves") +
  scale_x_date(breaks = pretty_breaks(n=10)) +
  scale_color_manual(values = c(rgb(5,61,96, maxColorValue = 255), rgb(255, 90, 52, maxColorValue = 255), "firebrick")) +
  theme(
    plot.background = element_rect(fill = rgb(242,242,242, maxColorValue = 255)),
    panel.background = element_rect(fill = "white"),
    panel.grid.major.y = element_line(size = 0.5, linetype = 'solid',colour = "gray"),
    panel.grid.major.x = element_line(size = 0.5, linetype = 'dashed',colour = "gray"),
    axis.title.x = element_text(colour = rgb(5,61,96, maxColorValue = 255), size = 16),
    axis.title.y = element_text(colour = rgb(5,61,96, maxColorValue = 255), size = 16),
    plot.title = element_text(colour = rgb(5,61,96, maxColorValue = 255), size = 14, face = "bold"),
    axis.text = element_text(size=16),
    text=element_text(family="Century Gothic"),
    legend.key = element_rect(fill = rgb(242,242,242, maxColorValue = 255)),
    legend.background=element_blank(),
    legend.title=element_blank(),
    legend.position="bottom"
  )


ggplotly(pic)%>%
  layout(legend = list(orientation = "h", x = 0.2, y = -0.2))



```




# Infrastructure {data-background="img/background_mtns_1.png"}

##  {data-background="img/flowSlide.PNG"}





# Core Strategies - Carry {data-background="img/background_mtns_1.png"}

## Carry overview {data-background="img/BlueBackgroundWithOrange.png"}

- Take advantage of curve shape
- Contango is typically less steep in further dated contracts, i.e. it has a lower roll yield than near dated contracts
- Extract roll yield with a short position in the near and a long position in the far dated contracts of the same commodity


## Carry Example {data-background="img/BlueBackgroundWithOrange.png"}



```{r}

plotdata <- contPrices[comdty == "LH" & tenure == "F3", .(index, `Front Month Price` = frontRP, `Deferred Month Price` = deferredRP)]
plotdata[, spread := `Deferred Month Price` - `Front Month Price`]
plotdata <- melt(plotdata, id.vars = "index")

plotdata[, year := year(index)]
plotdata[, month := month(index)]
plotdata[, md := max(index), by = .(year, month)]


ggplot(plotdata[index == md & variable != "spread" &  index >= "2000-01-01"], aes(x=index, y=value)) +
  geom_line(aes(group = variable, color = variable)) +
  ggtitle("Lean Hog Front nd Deferred Contract ETF prices") +
  theme_bw() +
  scale_y_continuous(breaks = pretty_breaks(n=10)) +
  scale_color_manual(values = c(rgb(255, 90, 52, maxColorValue = 255), rgb(5,61,96, maxColorValue = 255))) +
  theme(
    plot.background = element_rect(fill = rgb(242,242,242, maxColorValue = 255)),
    panel.background = element_rect(fill = "white"),
    panel.grid.major.y = element_line(size = 0.5, linetype = 'solid',colour = "gray"),
    panel.grid.major.x = element_line(size = 0.5, linetype = 'dashed',colour = "gray"),
    axis.title.x = element_text(colour = rgb(5,61,96, maxColorValue = 255), size = 16),
    axis.title.y = element_text(colour = rgb(5,61,96, maxColorValue = 255), size = 16),
    plot.title = element_text(colour = rgb(5,61,96, maxColorValue = 255), size = 18, face = "bold"),
    axis.text = element_text(size=16),
    text=element_text(family="Century Gothic"),
    legend.key = element_rect(fill = rgb(242,242,242, maxColorValue = 255)),
    legend.background=element_blank(),
    legend.title=element_blank(),
    legend.position="bottom"
  )




```


## Carry Example {data-background="img/BlueBackgroundWithOrange.png"}

```{r}

plotdata <- contPrices[comdty == "LH" & tenure == "F3" & index >= "2000-01-01", .(index, frontRP, deferredRP)]
plotdata <- melt(plotdata, id.vars = "index")

plotdata[, year := year(index)]
plotdata[, month := month(index)]


s.data <- split(plotdata, list(plotdata$year, plotdata$variable))

calculateRoll <- function(df){
  #df = s.data[[1]]
  start <- df[1, value]
  df[, value := value/start]
  df
}

yearData <- rbindlist(map(s.data, calculateRoll))

tempdata <- dcast(yearData, index ~ variable, value.var = "value")
tempdata[, value := deferredRP - frontRP]
tempdata <- tempdata[, .(index, value, variable = "spread")]
tempdata[, year := year(index)]
tempdata[, month := month(index)]


plotdata <- rbind(yearData, tempdata)

plotdata[, day := substring(index, 9, 10)]
plotdata[, date := paste(2000, month, day, sep="-")]
plotdata[, date := as.Date(date)]

meanSpreadData <- plotdata[variable == "spread", .(mean = mean(value), std = sd(value)), by = .(month)]
meanSpreadData <- meanSpreadData[order(month)]
  
ggplot(meanSpreadData, aes(x=month, y=mean)) +
  geom_line() +
  ylab("Average Carry") +
  xlab("Month") +
  geom_errorbar(aes(ymin=mean-std, ymax=mean+std)) +
  scale_x_continuous(breaks = pretty_breaks(n=10)) +
  scale_y_continuous(breaks = pretty_breaks(n=10)) +
  theme_bw() +
  ggtitle("Lean Hog Average Yearly Carry") +
  theme(
    plot.background = element_rect(fill = rgb(242,242,242, maxColorValue = 255)),
    panel.background = element_rect(fill = "white"),
    panel.grid.major.y = element_line(size = 0.5, linetype = 'solid',colour = "gray"),
    panel.grid.major.x = element_line(size = 0.5, linetype = 'dashed',colour = "gray"),
    axis.title.x = element_text(colour = rgb(5,61,96, maxColorValue = 255), size = 16),
    axis.title.y = element_text(colour = rgb(5,61,96, maxColorValue = 255), size = 16),
    plot.title = element_text(colour = rgb(5,61,96, maxColorValue = 255), size = 18, face = "bold"),
    axis.text = element_text(size=16),
    text=element_text(family="Century Gothic"),
    legend.key = element_rect(fill = rgb(242,242,242, maxColorValue = 255)),
    legend.background=element_blank(),
    legend.title=element_blank(),
    legend.position="bottom"
  )


```



## Flavours of carry {data-background="img/BlueBackgroundWithOrange.png"}

- **Carry - inspired by discretionary methodology (BB1)**
    - 30 commodities and 2 tenors
    - Sizing determined using percentile methodology
    - Risk 0.5% of strategy NAV per trade
    - Monthly rebalance

</br>

- **Machine Learning Enhanced Carry (BB2)**
    - 30 commodities and 2 tenors
    - Fractionally differenced time-series
    - Meta labelling
    - Ensemble machine learning
    - Deterine probability of profitable trade and size trade accordingly


## Backtest Assumptions {data-background="img/BlueBackgroundWithOrange.png"}

- $10 Round trip of each contract
- Slippage of 1 tick per contract

</br>

- Entry cost of one spread $= \text{\$}10 + 2 \text{ ticks} \times \text{multiplier}$
- Exit cost of one spread $= \text{\$}10 + 2 \text{ ticks} \times \text{multiplier}$


## BB1 - Statistics {data-background="img/BlueBackgroundWithOrange.png"}

```{r}

BB1returns <- as.data.table(read_csv('C:/Users/mauritz/Box Sync/Polar Star - Company Folder/1. Polar Star/Analysis/Analysis and Research/code/Quant Portfolio/data/BB1portfoioreturns.csv'))
BB1returns[!is.na(return), cumret := cumsum(return)]

BB1live <- as.data.table(read_csv("C:/Users/mauritz/Box Sync/Polar Star - Company Folder/1. Polar Star/Analysis/Analysis and Research/code/Quant Portfolio/data/BB1returns.csv"))
BB1live[, X1 := NULL]
setnames(BB1live, "index", "date")


BB1PSQCF <- as.data.table(read_csv("C:/Users/mauritz/Box Sync/Polar Star - Company Folder/1. Polar Star/Analysis/Analysis and Research/code/Quant Portfolio/data/Quant/BB1returns.csv"))
BB1PSQCF[, X1 := NULL]
setnames(BB1PSQCF, "index", "date")


```


```{r}

plotdata <- rbind(BB1returns[date < BB1live[, min(date)],.(date, return, type = "Backtest")], 
                  BB1live[date < BB1PSQCF[, min(date)],.(date, return = PnL, type = "Live")],
                  BB1PSQCF[, .(date, return = PnL, type = "PSQCF")])
plotdata[!is.na(return), cumret := cumsum(return)]

pic <- ggplot(plotdata, aes(x=date, y=cumret, group = type,
                            text = paste('Cummulative Return: ', round(cumret,2),
                                         '<br>Date: ', as.Date(date))
                            )
              ) +
  geom_line(aes(color = type)) +
  ylab("Cummulative Return") +
  xlab("") +
  ggtitle("BB1 Equity Curve") +
  scale_x_date(breaks = pretty_breaks(n = 10)) +
  scale_y_continuous(breaks = pretty_breaks(n = 10)) +
  scale_color_manual(values = c(rgb(255, 90, 52, maxColorValue = 255), rgb(5,61,96, maxColorValue = 255), "firebrick")) +
  theme(
    plot.background = element_rect(fill = rgb(242,242,242, maxColorValue = 255)),
    panel.background = element_rect(fill = "white"),
    panel.grid.major.y = element_line(size = 0.5, linetype = 'solid',colour = "gray"),
    panel.grid.major.x = element_line(size = 0.5, linetype = 'dashed',colour = "gray"),
    axis.title.x = element_text(colour = rgb(5,61,96, maxColorValue = 255), size = 16),
    axis.title.y = element_text(colour = rgb(5,61,96, maxColorValue = 255), size = 16),
    plot.title = element_text(colour = rgb(5,61,96, maxColorValue = 255), size = 18, face = "bold"),
    axis.text = element_text(size=16),
    text=element_text(family="Century Gothic"),
    legend.key = element_rect(fill = rgb(242,242,242, maxColorValue = 255)),
    legend.background=element_blank(),
    legend.title=element_blank(),
    legend.position="bottom"
  )


ggplotly(pic, tooltip = c("text")) %>%
  layout(legend = list(orientation = "h", x = 0.4, y = -0.2))

```


## BB1 - Statistics {data-background="img/BlueBackgroundWithOrange.png"}

```{r}

meanrollingsharpe <- BB1returns[, mean(`rolling return`, na.rm = TRUE)]


ggplot(BB1returns, aes(x=`rolling return`)) +
  geom_histogram(aes(y=..density..), fill = rgb(255, 90, 52, maxColorValue = 255)) +
  ylab("Probability Density") +
  xlab("Rolling Return (%)") +
  ggtitle("Histogram of BB1 12 month rolling return") +
  scale_x_continuous(breaks = pretty_breaks(n = 10)) +
  scale_y_continuous(breaks = pretty_breaks(n = 10)) +
  geom_vline(xintercept = meanrollingsharpe, color = rgb(5,61,96, maxColorValue = 255), linetype = "dashed", size = 1.5) +
  theme(
    plot.background = element_rect(fill = rgb(242,242,242, maxColorValue = 255)),
    panel.background = element_rect(fill = "white"),
    panel.grid.major.y = element_line(size = 0.5, linetype = 'solid',colour = "gray"),
    panel.grid.major.x = element_line(size = 0.5, linetype = 'dashed',colour = "gray"),
    axis.title.x = element_text(colour = rgb(5,61,96, maxColorValue = 255), size = 16),
    axis.title.y = element_text(colour = rgb(5,61,96, maxColorValue = 255), size = 16),
    plot.title = element_text(colour = rgb(5,61,96, maxColorValue = 255), size = 18, face = "bold"),
    axis.text = element_text(size=16),
    text=element_text(family="Century Gothic")
  )

```


## BB1 - Statistics {data-background="img/BlueBackgroundWithOrange.png"}

```{r}

meanrollingsharpe <- BB1returns[, mean(`rolling Sharpe`, na.rm = TRUE)]

ggplot(BB1returns, aes(x=`rolling Sharpe`)) +
  geom_histogram(aes(y=..density..), fill = rgb(255, 90, 52, maxColorValue = 255)) +
  ylab("Probability Density") +
  xlab("Rolling Sharpe Ratio") +
  ggtitle("Histogram of BB1 12 month rolling Sharpe Ratio") +
  scale_x_continuous(breaks = pretty_breaks(n = 10)) +
  scale_y_continuous(breaks = pretty_breaks(n = 10)) +
  geom_vline(xintercept = meanrollingsharpe, color = rgb(5,61,96, maxColorValue = 255), linetype = "dashed", size = 1.5) +
  theme(
    plot.background = element_rect(fill = rgb(242,242,242, maxColorValue = 255)),
    panel.background = element_rect(fill = "white"),
    panel.grid.major.y = element_line(size = 0.5, linetype = 'solid',colour = "gray"),
    panel.grid.major.x = element_line(size = 0.5, linetype = 'dashed',colour = "gray"),
    axis.title.x = element_text(colour = rgb(5,61,96, maxColorValue = 255), size = 16),
    axis.title.y = element_text(colour = rgb(5,61,96, maxColorValue = 255), size = 16),
    plot.title = element_text(colour = rgb(5,61,96, maxColorValue = 255), size = 18, face = "bold"),
    axis.text = element_text(size=16),
    text=element_text(family="Century Gothic")
  )

```






## BB2 - Statistics {data-background="img/BlueBackgroundWithOrange.png"}

```{r}

BB2returns <- as.data.table(read_csv('C:/Users/mauritz/Box Sync/Polar Star - Company Folder/1. Polar Star/Analysis/Analysis and Research/code/Quant Portfolio/data/calendarPortfolio_EWMA_RSI.csv'))
BB2returns[!is.na(return), cumret := cumsum(return)]

BB2live <- as.data.table(read_csv("C:/Users/mauritz/Box Sync/Polar Star - Company Folder/1. Polar Star/Analysis/Analysis and Research/code/Quant Portfolio/data/BB2returns.csv"))
BB2live[, X1 := NULL]
setnames(BB2live, "index", "date")


BB2PSQCF <- as.data.table(read_csv("C:/Users/mauritz/Box Sync/Polar Star - Company Folder/1. Polar Star/Analysis/Analysis and Research/code/Quant Portfolio/data/Quant/BB2returns.csv"))
BB2PSQCF[, X1 := NULL]
setnames(BB2PSQCF, "index", "date")


```


```{r}

plotdata <- rbind(BB2returns[date < BB2live[, min(date)],.(date, return, type = "Backtest")], 
                  BB2live[date < BB2PSQCF[, min(date)],.(date, return = PnL, type = "Live")],
                  BB2PSQCF[, .(date, return = PnL, type = "PSQCF")])
plotdata[!is.na(return), cumret := cumsum(return)]

pic <- ggplot(plotdata, aes(x=date, y=cumret, group = type,
                            text = paste('Cummulative Return: ', round(cumret,2),
                                         '<br>Date: ', as.Date(date))
                            )
              ) +
  geom_line(aes(color = type)) +
  ylab("Cummulative Return") +
  xlab("") +
  ggtitle("BB2 Equity Curve") +
  scale_x_date(breaks = pretty_breaks(n = 10)) +
  scale_y_continuous(breaks = pretty_breaks(n = 10)) +
  scale_color_manual(values = c(rgb(255, 90, 52, maxColorValue = 255), rgb(5,61,96, maxColorValue = 255), "firebrick")) +
  theme(
    plot.background = element_rect(fill = rgb(242,242,242, maxColorValue = 255)),
    panel.background = element_rect(fill = "white"),
    panel.grid.major.y = element_line(size = 0.5, linetype = 'solid',colour = "gray"),
    panel.grid.major.x = element_line(size = 0.5, linetype = 'dashed',colour = "gray"),
    axis.title.x = element_text(colour = rgb(5,61,96, maxColorValue = 255), size = 16),
    axis.title.y = element_text(colour = rgb(5,61,96, maxColorValue = 255), size = 16),
    plot.title = element_text(colour = rgb(5,61,96, maxColorValue = 255), size = 18, face = "bold"),
    axis.text = element_text(size=16),
    text=element_text(family="Century Gothic"),
    legend.key = element_rect(fill = rgb(242,242,242, maxColorValue = 255)),
    legend.background=element_blank(),
    legend.title=element_blank(),
    legend.position="bottom"
  )


ggplotly(pic, tooltip = c("text")) %>%
  layout(legend = list(orientation = "h", x = 0.4, y = -0.2))

```

## BB2 - Statistics {data-background="img/BlueBackgroundWithOrange.png"}

```{r}

meanrollingsharpe <- BB2returns[, mean(`rolling return`, na.rm = TRUE)]


ggplot(BB2returns, aes(x=`rolling return`)) +
  geom_histogram(aes(y=..density..), fill = rgb(255, 90, 52, maxColorValue = 255)) +
  ylab("Probability Density") +
  xlab("Rolling Return (%)") +
  ggtitle("Histogram of BB2 12 month rolling return") +
  scale_x_continuous(breaks = pretty_breaks(n = 10)) +
  scale_y_continuous(breaks = pretty_breaks(n = 10)) +
  geom_vline(xintercept = meanrollingsharpe, color = rgb(5,61,96, maxColorValue = 255), linetype = "dashed", size = 1.5) +
  theme(
    plot.background = element_rect(fill = rgb(242,242,242, maxColorValue = 255)),
    panel.background = element_rect(fill = "white"),
    panel.grid.major.y = element_line(size = 0.5, linetype = 'solid',colour = "gray"),
    panel.grid.major.x = element_line(size = 0.5, linetype = 'dashed',colour = "gray"),
    axis.title.x = element_text(colour = rgb(5,61,96, maxColorValue = 255), size = 16),
    axis.title.y = element_text(colour = rgb(5,61,96, maxColorValue = 255), size = 16),
    plot.title = element_text(colour = rgb(5,61,96, maxColorValue = 255), size = 18, face = "bold"),
    axis.text = element_text(size=16),
    text=element_text(family="Century Gothic")
  )

```


## BB2 - Statistics {data-background="img/BlueBackgroundWithOrange.png"}

```{r}

meanrollingsharpe <- BB2returns[, mean(`rolling Sharpe`, na.rm = TRUE)]

ggplot(BB2returns, aes(x=`rolling Sharpe`)) +
  geom_histogram(aes(y=..density..), fill = rgb(255, 90, 52, maxColorValue = 255)) +
  ylab("Probability Density") +
  xlab("Rolling Sharpe Ratio") +
  ggtitle("Histogram of BB2 12 month rolling Sharpe Ratio") +
  scale_x_continuous(breaks = pretty_breaks(n = 10)) +
  scale_y_continuous(breaks = pretty_breaks(n = 10)) +
  geom_vline(xintercept = meanrollingsharpe, color = rgb(5,61,96, maxColorValue = 255), linetype = "dashed", size = 1.5) +
  theme(
    plot.background = element_rect(fill = rgb(242,242,242, maxColorValue = 255)),
    panel.background = element_rect(fill = "white"),
    panel.grid.major.y = element_line(size = 0.5, linetype = 'solid',colour = "gray"),
    panel.grid.major.x = element_line(size = 0.5, linetype = 'dashed',colour = "gray"),
    axis.title.x = element_text(colour = rgb(5,61,96, maxColorValue = 255), size = 16),
    axis.title.y = element_text(colour = rgb(5,61,96, maxColorValue = 255), size = 16),
    plot.title = element_text(colour = rgb(5,61,96, maxColorValue = 255), size = 18, face = "bold"),
    axis.text = element_text(size=16),
    text=element_text(family="Century Gothic")
  )

```



## Systematic Carry - Statistics {data-background="img/BlueBackgroundWithOrange.png"}

```{r}

tbl1 <- table.AnnualizedReturns(BB1returns[, .(date, BB1 = return)])
nms <- row.names(tbl1)
tbl1 <- as.data.table(tbl1)
tbl1[, variable := nms]

BB1returns[, month := month(date)]
BB1returns[, year := year(date)]
BB1returns[, md := min(date), by = .(year, month)]

#monthlydata <- CalculateReturns(BB1returns[date == md,.(date, normalised)])
#monthlydata <- as.data.table(monthlydata)
monthlydata <- BB1returns[,.(normalised = prod(return + 1) - 1), by = .(year, month)]
monthlydata[, sign := sign(normalised)]

avgneg <- monthlydata[sign == -1, mean(normalised) * 100]
avgpos <- monthlydata[sign == 1, mean(normalised) * 100]

noneg <- nrow(monthlydata[sign == -1])
nopos <- nrow(monthlydata[sign == 1])


maxdraw <- maxDrawdown(as.xts(BB1returns[, .(date, return)]))[1] * 100
ret.maxdraw <- maxdraw/tbl1[variable == "Annualized Return", BB1 * 100]

parttoadd <- data.table(
  variable = c("Maximum Drawdown", "Maximum Drawdown/Annualized Return", "Number of Positive Months", "Number of Negative Months", "Average Positive Month Return", "Average Negative Month Return"),
  BB1 = c(maxdraw, ret.maxdraw, nopos, noneg, avgpos, avgneg)
)


tbl1[variable != "Annualized Sharpe (Rf=0%)", BB1 := BB1 * 100]

tbl1 <- rbind(tbl1, parttoadd)
tbl1 <- tbl1[,.(variable, BB1)]
tbl1[, BB1 := round(BB1, 3)]

setnames(tbl1, "variable", "Statistic")

##

BB1tabledata <- rbind(BB1live[date < BB1PSQCF[, min(date)], .(date, `BB1 live` = PnL)], BB1PSQCF[,.(date, `BB1 live` = PnL)])

tbl1live <- table.AnnualizedReturns(BB1tabledata)
nms <- row.names(tbl1live)
tbl1live <- as.data.table(tbl1live)
tbl1live[, variable := nms]


BB1tabledata[, month := month(date)]
BB1tabledata[, year := year(date)]
BB1tabledata[, md := min(date), by = .(year, month)]

BB1tabledata[is.na(`BB1 live`), `BB1 live` := 0]
BB1tabledata[, normalised := cumprod(`BB1 live` + 1)]

#monthlydata <- CalculateReturns(BB1tabledata[date == md,.(date, normalised)])
monthlydata <- BB1tabledata[,.(normalised = prod(`BB1 live` + 1) - 1), by = .(year, month)]
#monthlydata <- as.data.table(monthlydata)
monthlydata[, sign := sign(normalised)]

avgneg <- monthlydata[sign == -1, mean(normalised) * 100]
avgpos <- monthlydata[sign == 1, mean(normalised) * 100]

noneg <- nrow(monthlydata[sign == -1])
nopos <- nrow(monthlydata[sign == 1])


maxdraw <- maxDrawdown(as.xts(BB1tabledata[, .(date, `BB1 live`)]))[1] * 100
ret.maxdraw <- maxdraw/tbl1live[variable == "Annualized Return", `BB1 live` * 100]

parttoadd <- data.table(
  variable = c("Maximum Drawdown", "Maximum Drawdown/Annualized Return", "Number of Positive Months", "Number of Negative Months", "Average Positive Month Return", "Average Negative Month Return"),
  `BB1 live` = c(maxdraw, ret.maxdraw, nopos, noneg, avgpos, avgneg)
)


tbl1live[variable != "Annualized Sharpe (Rf=0%)", `BB1 live` := `BB1 live` * 100]

tbl1live <- rbind(tbl1live, parttoadd)
tbl1live <- tbl1live[,.(variable, `BB1 live`)]
tbl1live[, `BB1 live` := round(`BB1 live`, 3)]

setnames(tbl1live, "variable", "Statistic")


##


tbl2 <- table.AnnualizedReturns(BB2returns[, .(date, BB2 = return)])
nms <- row.names(tbl2)
tbl2 <- as.data.table(tbl2)
tbl2[, variable := nms]

tbl2[variable != "Annualized Sharpe (Rf=0%)", BB2 := BB2 * 100]

BB2returns[, month := month(date)]
BB2returns[, year := year(date)]
BB2returns[, md := min(date), by = .(year, month)]

#monthlydata <- CalculateReturns(BB2returns[date == md,.(date, normalised)])
#monthlydata <- as.data.table(monthlydata)
monthlydata <- BB1returns[,.(normalised = prod(return + 1) - 1), by = .(year, month)]
monthlydata[, sign := sign(normalised)]

avgneg <- monthlydata[sign == -1, mean(normalised) * 100]
avgpos <- monthlydata[sign == 1, mean(normalised) * 100]

noneg <- nrow(monthlydata[sign == -1])
nopos <- nrow(monthlydata[sign == 1])


maxdraw <- maxDrawdown(as.xts(BB2returns[, .(date, return)]))[1] * 100
ret.maxdraw <- maxdraw/tbl2[variable == "Annualized Return", BB2 ]

parttoadd <- data.table(
  variable = c("Maximum Drawdown", "Maximum Drawdown/Annualized Return", "Number of Positive Months", "Number of Negative Months", "Average Positive Month Return", "Average Negative Month Return"),
  BB2 = c(maxdraw, ret.maxdraw, nopos, noneg, avgpos, avgneg)
)


tbl2 <- rbind(tbl2, parttoadd)
tbl2 <- tbl2[,.(variable, BB2)]
tbl2[, BB2 := round(BB2, 3)]

setnames(tbl2, "variable", "Statistic")


##

BB2tabledata <- rbind(BB2live[date < BB2PSQCF[, min(date)], .(date, `BB2 live` = PnL)], BB2PSQCF[, .(date, `BB2 live` = PnL)])

tbl2live <- table.AnnualizedReturns(BB2tabledata)
nms <- row.names(tbl2live)
tbl2live <- as.data.table(tbl2live)
tbl2live[, variable := nms]

BB2tabledata[, month := month(date)]
BB2tabledata[, year := year(date)]
BB2tabledata[, md := min(date), by = .(year, month)]

BB2tabledata[is.na(`BB2 live`), `BB2 live` := 0]
BB2tabledata[, normalised := cumprod(`BB2 live` + 1)]

#monthlydata <- CalculateReturns(BB2tabledata[date == md,.(date, normalised)])
#monthlydata <- as.data.table(monthlydata)
monthlydata <- BB2tabledata[,.(normalised = prod(`BB2 live` + 1) - 1), by = .(year, month)]
monthlydata[, sign := sign(normalised)]

avgneg <- monthlydata[sign == -1, mean(normalised) * 100]
avgpos <- monthlydata[sign == 1, mean(normalised) * 100]

noneg <- nrow(monthlydata[sign == -1])
nopos <- nrow(monthlydata[sign == 1])


maxdraw <- maxDrawdown(as.xts(BB2tabledata[, .(date, `BB2 live`)]))[1] * 100
ret.maxdraw <- maxdraw/tbl2live[variable == "Annualized Return", `BB2 live` * 100]

parttoadd <- data.table(
  variable = c("Maximum Drawdown", "Maximum Drawdown/Annualized Return", "Number of Positive Months", "Number of Negative Months", "Average Positive Month Return", "Average Negative Month Return"),
  `BB2 live` = c(maxdraw, ret.maxdraw, nopos, noneg, avgpos, avgneg)
)


tbl2live[variable != "Annualized Sharpe (Rf=0%)", `BB2 live` := `BB2 live` * 100]

tbl2live <- rbind(tbl2live, parttoadd)
tbl2live <- tbl2live[,.(variable, `BB2 live`)]
tbl2live[, `BB2 live` := round(`BB2 live`, 3)]

setnames(tbl2live, "variable", "Statistic")


##

tblA <- merge(tbl1, tbl1live, by = "Statistic")
tblB <- merge(tbl2, tbl2live, by = "Statistic")
tbl <- merge(tblA, tblB, by = "Statistic")

formattable(tbl)

```



## Carry literature {data-background="img/BlueBackgroundWithOrange.png"}

Literature on extracting carry from futures:

- <a href="http://onlinelibrary.wiley.com/doi/10.1111/jofi.12096/full" target="_blank">An Anatomy of Commodity Futures Risk Premia - Szymanowska et al.</a>
- <a href="http://www.sciencedirect.com/science/article/pii/S0140988305001027" target="_blank">Commodity convenience yield and risk premium determination - Wei and Zhu</a>
- <a href="https://papers.ssrn.com/sol3/papers.cfm?abstract_id=2298565" target="_blank">Carry - Koijen et al.</a>
- <a href = "https://faculty.fuqua.duke.edu/~charvey/Research/Published_Papers/P91_The_strategic_and.pdf">The Strategic and Tactical Value of Commodity Futures - Erb and Harvey</a>


Literature on applying machine learing techniques in algorithmic trading:

- <a href = "https://www.amazon.com/Advances-Financial-Machine-Learning-Marcos-ebook/dp/B079KLDW21" target="_blank">Advanced in financial machine learning - Marcos López de Prado</a>
- <a href = "https://papers.ssrn.com/sol3/papers.cfm?abstract_id=3104816" target="_blank">The 10 Reasons Most Machine Learning Funds Fail</a>




# Core Strategies - Trend  {data-background="img/background_mtns_1.png"}

## Trend overview {data-background="img/BlueBackgroundWithOrange.png"}

- Trend following is about absolute performance of each commodity
- Identify trends over selection of time frames
- Slowly build position as trend increases
- Slowly exit position as trend decreases
- 35 Commodities and 2 tenors


## Designing, not fitting a strategy {data-background="img/BlueBackgroundWithOrange.png"}

Aim of a trend following strategy

- Profitably trade trending markets
- Accross a diverse universe of commodities

</br>

If we feed our trend system fake trendy data

- linear and sinusoidal, with 
- Gaussian noise
 
can we trade it profitably?


## Fake Trendy Data {data-background="img/BlueBackgroundWithOrange.png"}


```{r}

allPrices <- as.data.table(read_csv("C:/Users/mauritz/Box Sync/Polar Star - Company Folder/1. Polar Star/Analysis/Analysis and Research/code/Quant Start Ideas/PS trend/designing strategies/allPrices.csv"))
setnames(allPrices, names(allPrices), c("date", "price", "comdty"))

allPrices[, ref := substring(comdty, 3, 3)]
allPrices[, comdty := substring(comdty, 1, 2)]


```


```{r}

ggplot(allPrices[date >= "2003-01-01" & ref > 3], aes(x=date, y=price)) +
  geom_line() +
  facet_grid(ref ~ comdty) +
  ggtitle("Time evolution of the fake data") +
  xlab("") +
  ylab("Price") +
  scale_y_continuous(breaks = pretty_breaks(n = 8)) +
  theme(
    plot.background = element_rect(fill = rgb(242,242,242, maxColorValue = 255)),
    panel.background = element_rect(fill = "white"),
    panel.grid.major.y = element_line(size = 0.5, linetype = 'solid',colour = "gray"),
    panel.grid.major.x = element_line(size = 0.5, linetype = 'dashed',colour = "gray"),
    axis.title.x = element_text(colour = rgb(5,61,96, maxColorValue = 255), size = 16),
    axis.title.y = element_text(colour = rgb(5,61,96, maxColorValue = 255), size = 16),
    plot.title = element_text(colour = rgb(5,61,96, maxColorValue = 255), size = 18, face = "bold"),
    axis.text = element_text(size=16),
    text=element_text(family="Century Gothic")
  )


```

## Trend - Fake Data Performance {data-background="img/BlueBackgroundWithOrange.png"}

```{r}

returns_instrument_forecast_divmult_scale_weight_estimates <- as.data.table(read_csv("C:/Users/mauritz/Box Sync/Polar Star - Company Folder/1. Polar Star/Analysis/Analysis and Research/code/Quant Start Ideas/PS trend/designing strategies/returns_instrument_forecast_divmult_scale_weight_estimates.csv"))
setnames(returns_instrument_forecast_divmult_scale_weight_estimates, names(returns_instrument_forecast_divmult_scale_weight_estimates), c("date", "return"))

returns_instrument_forecast_divmult_scale_weight_estimates <- returns_instrument_forecast_divmult_scale_weight_estimates[date >= "2002-12-31"]
returns_instrument_forecast_divmult_scale_weight_estimates[, return := return/100]
returns_instrument_forecast_divmult_scale_weight_estimates[, cumret := cumsum(return)]

```

```{r}

ggplot(returns_instrument_forecast_divmult_scale_weight_estimates, aes(x=date, y=cumret)) +
  geom_line(colour = rgb(5,61,96, maxColorValue = 255)) +
  ggtitle("Cummulative performance on fake trendy data") +
  xlab("") +
  ylab("P&L") +
  scale_y_continuous(breaks = pretty_breaks(n = 8)) +
  theme(
    plot.background = element_rect(fill = rgb(242,242,242, maxColorValue = 255)),
    panel.background = element_rect(fill = "white"),
    panel.grid.major.y = element_line(size = 0.5, linetype = 'solid',colour = "gray"),
    panel.grid.major.x = element_line(size = 0.5, linetype = 'dashed',colour = "gray"),
    axis.title.x = element_text(colour = rgb(5,61,96, maxColorValue = 255), size = 16),
    axis.title.y = element_text(colour = rgb(5,61,96, maxColorValue = 255), size = 16),
    plot.title = element_text(colour = rgb(5,61,96, maxColorValue = 255), size = 18, face = "bold"),
    axis.text = element_text(size=16),
    text=element_text(family="Century Gothic")
  )



```

## Trend - Statistics {data-background="img/BlueBackgroundWithOrange.png"}

```{r}

TR1returns <- as.data.table(read_csv('C:/Users/mauritz/Box Sync/Polar Star - Company Folder/1. Polar Star/Analysis/Analysis and Research/code/Quant Portfolio/data/trendPortfolioReturns.csv'))
TR1returns <- TR1returns[ , .(date = X1, return = `percent return`/100)]
TR1returns <- TR1returns[date >= "1998-01-02"]
TR1returns[!is.na(return), cumret := cumsum(return)]


TR1returns[, normalised := cumprod(1 + return)]
startval <- TR1returns[1, normalised]
TR1returns[, normalised := normalised/startval]

TR1returns[, prevnorm := guyrot(normalised, 252)]
TR1returns[1:252, prevnorm := NA]

TR1returns[, `rolling return` := (normalised - prevnorm)/prevnorm * 100] 

rllsd <- rollapply(TR1returns[, return], width = 252, FUN = sd)
rllsd <- c(rep(NA, 251), rllsd)

TR1returns[, `rolling risk` := rllsd * sqrt(252) * 100]
TR1returns[, `rolling Sharpe` := `rolling return`/`rolling risk`]




TR1live <- as.data.table(read_csv("C:/Users/mauritz/Box Sync/Polar Star - Company Folder/1. Polar Star/Analysis/Analysis and Research/code/Quant Portfolio/data/TR1returns.csv"))
TR1live[, X1 := NULL]
setnames(TR1live, "index", "date")


TR1PSQCF <- as.data.table(read_csv("C:/Users/mauritz/Box Sync/Polar Star - Company Folder/1. Polar Star/Analysis/Analysis and Research/code/Quant Portfolio/data/Quant/TR1returns.csv"))
TR1PSQCF[, X1 := NULL]
setnames(TR1PSQCF, "index", "date")



```


```{r}

plotdata <- rbind(TR1returns[date < TR1live[, min(date)],.(date, return, type = "Backtest")], 
                  TR1live[date < TR1PSQCF[, min(date)],.(date, return = PnL, type = "Live")],
                  TR1PSQCF[, .(date, return = PnL, type = "PSQCF")])
plotdata[!is.na(return), cumret := cumsum(return)]

pic <- ggplot(plotdata, aes(x=date, y=cumret, group = type,
                            text = paste('Cummulative Return: ', round(cumret,2),
                                         '<br>Date: ', as.Date(date))
                            )
              ) +
  geom_line(aes(color = type)) +
  ylab("Cummulative Return") +
  xlab("") +
  ggtitle("TR1 Equity Curve") +
  scale_x_date(breaks = pretty_breaks(n = 10)) +
  scale_y_continuous(breaks = pretty_breaks(n = 10)) +
  scale_color_manual(values = c(rgb(255, 90, 52, maxColorValue = 255), rgb(5,61,96, maxColorValue = 255), "firebrick")) +
  theme(
    plot.background = element_rect(fill = rgb(242,242,242, maxColorValue = 255)),
    panel.background = element_rect(fill = "white"),
    panel.grid.major.y = element_line(size = 0.5, linetype = 'solid',colour = "gray"),
    panel.grid.major.x = element_line(size = 0.5, linetype = 'dashed',colour = "gray"),
    axis.title.x = element_text(colour = rgb(5,61,96, maxColorValue = 255), size = 16),
    axis.title.y = element_text(colour = rgb(5,61,96, maxColorValue = 255), size = 16),
    plot.title = element_text(colour = rgb(5,61,96, maxColorValue = 255), size = 18, face = "bold"),
    axis.text = element_text(size=16),
    text=element_text(family="Century Gothic"),
    legend.key = element_rect(fill = rgb(242,242,242, maxColorValue = 255)),
    legend.background=element_blank(),
    legend.title=element_blank(),
    legend.position="bottom"
  )


ggplotly(pic, tooltip = c("text")) %>%
  layout(legend = list(orientation = "h", x = 0.4, y = -0.2))

```

## Trend - Statistics {data-background="img/BlueBackgroundWithOrange.png"}

```{r}

meanrollingsharpe <- TR1returns[, mean(`rolling return`, na.rm = TRUE)]


ggplot(TR1returns, aes(x=`rolling return`)) +
  geom_histogram(aes(y=..density..), fill = rgb(255, 90, 52, maxColorValue = 255)) +
  ylab("Probability Density") +
  xlab("Rolling Return (%)") +
  ggtitle("Histogram of TR1 12 month rolling return") +
  scale_x_continuous(breaks = pretty_breaks(n = 10)) +
  scale_y_continuous(breaks = pretty_breaks(n = 10)) +
  geom_vline(xintercept = meanrollingsharpe, color = rgb(5,61,96, maxColorValue = 255), linetype = "dashed", size = 1.5) +
  theme(
    plot.background = element_rect(fill = rgb(242,242,242, maxColorValue = 255)),
    panel.background = element_rect(fill = "white"),
    panel.grid.major.y = element_line(size = 0.5, linetype = 'solid',colour = "gray"),
    panel.grid.major.x = element_line(size = 0.5, linetype = 'dashed',colour = "gray"),
    axis.title.x = element_text(colour = rgb(5,61,96, maxColorValue = 255), size = 16),
    axis.title.y = element_text(colour = rgb(5,61,96, maxColorValue = 255), size = 16),
    plot.title = element_text(colour = rgb(5,61,96, maxColorValue = 255), size = 18, face = "bold"),
    axis.text = element_text(size=16),
    text=element_text(family="Century Gothic")
  )

```


## Trend - Statistics {data-background="img/BlueBackgroundWithOrange.png"}

```{r}

meanrollingsharpe <- TR1returns[, mean(`rolling Sharpe`, na.rm = TRUE)]

ggplot(TR1returns, aes(x=`rolling Sharpe`)) +
  geom_histogram(aes(y=..density..), fill = rgb(255, 90, 52, maxColorValue = 255)) +
  ylab("Probability Density") +
  xlab("Rolling Sharpe Ratio") +
  ggtitle("Histogram of TR1 12 month rolling Sharpe Ratio") +
  scale_x_continuous(breaks = pretty_breaks(n = 10)) +
  scale_y_continuous(breaks = pretty_breaks(n = 10)) +
  geom_vline(xintercept = meanrollingsharpe, color = rgb(5,61,96, maxColorValue = 255), linetype = "dashed", size = 1.5) +
  theme(
    plot.background = element_rect(fill = rgb(242,242,242, maxColorValue = 255)),
    panel.background = element_rect(fill = "white"),
    panel.grid.major.y = element_line(size = 0.5, linetype = 'solid',colour = "gray"),
    panel.grid.major.x = element_line(size = 0.5, linetype = 'dashed',colour = "gray"),
    axis.title.x = element_text(colour = rgb(5,61,96, maxColorValue = 255), size = 16),
    axis.title.y = element_text(colour = rgb(5,61,96, maxColorValue = 255), size = 16),
    plot.title = element_text(colour = rgb(5,61,96, maxColorValue = 255), size = 18, face = "bold"),
    axis.text = element_text(size=16),
    text=element_text(family="Century Gothic")
  )

```


## Trend - Statistics {data-background="img/BlueBackgroundWithOrange.png"}

```{r}

tbl1 <- table.AnnualizedReturns(TR1returns[, .(date, TR1 = return)])
nms <- row.names(tbl1)
tbl1 <- as.data.table(tbl1)
tbl1[, variable := nms]

tbl1[variable != "Annualized Sharpe (Rf=0%)", TR1 := TR1 * 100]

TR1returns[, month := month(date)]
TR1returns[, year := year(date)]
TR1returns[, md := min(date), by = .(year, month)]

#monthlydata <- CalculateReturns(TR1returns[date == md,.(date, normalised)])
#monthlydata <- as.data.table(monthlydata)
monthlydata <- TR1returns[,.(normalised = prod(return + 1) - 1), by = .(year, month)]
monthlydata[, sign := sign(normalised)]

avgneg <- monthlydata[sign == -1, mean(normalised) * 100]
avgpos <- monthlydata[sign == 1, mean(normalised * 100)]

noneg <- nrow(monthlydata[sign == -1])
nopos <- nrow(monthlydata[sign == 1])


maxdraw <- maxDrawdown(as.xts(BB1returns[, .(date, return)]))[1] * 100
ret.maxdraw <- maxdraw/tbl1[variable == "Annualized Return", TR1 ]

parttoadd <- data.table(
  variable = c("Maximum Drawdown", "Maximum Drawdown/Annualized Return", "Number of Positive Months", "Number of Negative Months", "Average Positive Month Return", "Average Negative Month Return"),
  TR1 = c(maxdraw, ret.maxdraw, nopos, noneg, avgpos, avgneg)
)


tbl1 <- rbind(tbl1, parttoadd)
tbl1 <- tbl1[,.(variable, TR1)]
tbl1[, TR1 := round(TR1, 3)]

##

TR1tabledata <- rbind(TR1live[date < TR1PSQCF[, min(date)], .(date, `TR1 live` = PnL)], TR1PSQCF[, .(date, `TR1 live` = PnL)])

tbl1live <- table.AnnualizedReturns(TR1tabledata)
nms <- row.names(tbl1live)
tbl1live <- as.data.table(tbl1live)
tbl1live[, variable := nms]

TR1tabledata[, month := month(date)]
TR1tabledata[, year := year(date)]
TR1tabledata[, md := min(date), by = .(year, month)]

TR1tabledata[is.na(`TR1 live`), `TR1 live` := 0]
TR1tabledata[, normalised := cumprod(`TR1 live` + 1)]

#monthlydata <- CalculateReturns(TR1tabledata[date == md,.(date, normalised)])
#monthlydata <- as.data.table(monthlydata)
monthlydata <- TR1tabledata[,.(normalised = prod(`TR1 live` + 1) - 1), by = .(year, month)]
monthlydata[, sign := sign(normalised)]

avgneg <- monthlydata[sign == -1, mean(normalised) * 100]
avgpos <- monthlydata[sign == 1, mean(normalised) * 100]

noneg <- nrow(monthlydata[sign == -1])
nopos <- nrow(monthlydata[sign == 1])


maxdraw <- maxDrawdown(as.xts(TR1tabledata[, .(date, `TR1 live`)]))[1] * 100
ret.maxdraw <- maxdraw/tbl1live[variable == "Annualized Return", `TR1 live` * 100]

parttoadd <- data.table(
  variable = c("Maximum Drawdown", "Maximum Drawdown/Annualized Return", "Number of Positive Months", "Number of Negative Months", "Average Positive Month Return", "Average Negative Month Return"),
  `TR1 live` = c(maxdraw, ret.maxdraw, nopos, noneg, avgpos, avgneg)
)


tbl1live[variable != "Annualized Sharpe (Rf=0%)", `TR1 live` := `TR1 live` * 100]

tbl1live <- rbind(tbl1live, parttoadd)
tbl1live <- tbl1live[,.(variable, `TR1 live`)]
tbl1live[, `TR1 live` := round(`TR1 live`, 3)]

setnames(tbl1live, "variable", "Statistic")


##

tbl <- cbind(tbl1, tbl1live[, .(`TR1 live`)])


formattable(tbl)

```



## Trend literature {data-background="img/BlueBackgroundWithOrange.png"}

- <a href = "https://faculty.fuqua.duke.edu/~charvey/Research/Published_Papers/P91_The_strategic_and.pdf" target="_blank">The Strategic and Tactical Value of Commodity Futures - Erb and Harvey</a>
- <a href = "http://faculty.som.yale.edu/garygorton/documents/fundamentals_Jan2012_FINAL.pdf" target="_blank">The fundamentals of Commodity Futures Returns</a>
- <a href = "http://pages.stern.nyu.edu/~lpederse/papers/TimeSeriesMomentum.pdf" target="_blank">Time series momentum - Moskowitz et al.</a>
- <a href = "https://papers.ssrn.com/sol3/papers.cfm?abstract_id=2603731" target="_blank">Which trend is your friend - Levine and Pedersen</a>
- <a href = "https://www.amazon.com/Trend-Following-Managed-Futures-Trading/dp/1118890973" target="_blank">Trend Following with Managed Futures - Greyserman and Kaminski</a>


In-house Research:

- <a href = "https://clever-stonebraker-da1d54.netlify.com/post/designing-a-trend-strategy/" target="_blank">Designing a trend strategy</a>





# Core Strategies - Relative Roll {data-background="img/background_mtns_1.png"}

## Relative Roll overview {data-background="img/BlueBackgroundWithOrange.png"}

Strategy not yet live.


## Relative Roll literature {data-background="img/BlueBackgroundWithOrange.png"}

- <a href = "https://www.nber.org/papers/w7032.pdf" target="_blank">Pairs Trading: Performance of a Relative Value Arbitrage Rule - Gatev et al.</a>
- <a href = "https://books.google.co.uk/books?hl=en&lr=&id=yK2mvtpcumcC&oi=fnd&pg=PR5&dq=Pairs+Trading:+Quantitative+Methods+and+Analysis&ots=bUwnOD9XTA&sig=604hSkiMkxh7kzqtsG_Ze2w_eno&redir_esc=y" target="_blank">Pairs Trading - Vidyamurthy</a>
- <a href = "http://statistik.econ.kit.edu/download/doc_secure1/ptem342v144techreport.pdf">The Application of Pairs Trading to Energy Futures Markets - Kanamura et al.</a>




# Core Strategies - Sentiment {data-background="img/background_mtns_1.png"}

## Sentiment overview {data-background="img/BlueBackgroundWithOrange.png"}

Strategy not yet live.



## Sentiment literature {data-background="img/BlueBackgroundWithOrange.png"}

- <a href = "https://www.ravenpack.com/" target="_blank">Machine learning and event detection for trading energy and metals futures - Hafez et al.</a>
- <a href = "https://www.ravenpack.com/" target="_blank">JP Morgan:  Using Ravenpack sentiment for cross asset style timing - Hafez et al.</a>
- <a href = "https://efmaefm.org/0EFMAMEETINGS/EFMA%20ANNUAL%20MEETINGS/2017-Athens/papers/EFMA2017_0513_fullpaper.pdf" target="_blank">Macro Fundamentals or Geopolitical Events? A Textual Analysis of News Events for Crude Oil - Brandt and Gao</a>




# Quantitative Portfolio {data-background="img/background_mtns_1.png"}

## Investment Thesis {data-background="img/BlueBackgroundWithOrange.png"}

- Focus on the intersection of technolody, data and behavioral finance applied to the broad commodity space
- Build strategies based on sound economic theory to help deliver long-term repeatable results.  Inspired by
    - academic and
    - proprietary research.
- Investment process built on the scientific method consisting of the systematic
    - observation,
    - measurement, 
    - experiment,
    - hypothesis formulation,
    - testing and
    - modification of hypothesis


## Quantitative Portfolio - Statistics {data-background="img/BlueBackgroundWithOrange.png"}

```{r}

bb1 <- BB1returns[, .(date, BB1 = return)]
bb2 <- BB2returns[, .(date, BB2 = return)]
tr1 <- TR1returns[, .(date, TR1 = return)]

returns <- merge(bb1, bb2, by = "date", all = TRUE)
returns <- merge(returns, tr1, by = "date", all = TRUE)

#returns[, portfolio := BB1 * 0.3 + BB2 * 0.5 + TR1 * 0.2]
returns[, portfolio := BB1 * 0 + BB2 * 0.5 + TR1 * 0.5]

## perform a couple calcs
portfolio <- returns[ , .(date, return = portfolio)]

portfolio <- portfolio[!is.na(return)]


##... FEES ...##

portfolio[, year := year(date)]
portfolio[, month := month(date)]

portfolio[, md := max(date), by = .(year, month)]

portfolio[, management := 0]

portfolio[, half := ifelse(month < 7, 1, 2)]


portfolio[date == md, management := 0.02/12]

portfolio[, totalreturn := return - management]


s.data <- split(portfolio, list(portfolio$year, portfolio$half))
tflist <- map_lgl(s.data, function(df){ifelse(nrow(df) < 1, F, T)})
s.data <- s.data[tflist]

getPerformance <- function(df, hurdle = 1.025){
  #df = s.data[[1]]
  ret <- df[, prod(1 + totalreturn)]
  diff <- ret - hurdle
  perf <- ifelse(diff > 0, ret - 1.02, 0) * 0.2
  
  output <- data.table(
    date = df[, max(date)],
    performance = perf
  )
}

performancefees <- rbindlist(lapply(s.data, getPerformance))

portfolio <- merge(portfolio, performancefees, by = "date", all = TRUE)
portfolio[, performance := ifelse(is.na(performance), 0, performance)]

portfolio[, totalreturn := totalreturn - performance]

##...      ...##


portfolio[, cumret := cumsum(totalreturn)]
portfolio[, normalised := cumprod(1 + totalreturn)]
startval <- portfolio[1, normalised]
portfolio[, normalised := normalised/startval]



portfolio[, prevnorm := guyrot(normalised, 252)]
portfolio[1:252, prevnorm := NA]

portfolio[, `rolling return` := (normalised - prevnorm)/prevnorm * 100] 

rllsd <- rollapply(portfolio[, return], width = 252, FUN = sd)
rllsd <- c(rep(NA, 251), rllsd)

portfolio[, `rolling risk` := rllsd * sqrt(252) * 100]
portfolio[, `rolling Sharpe` := `rolling return`/`rolling risk`]


```


```{r}

# bb1live <- BB1live[, .(date, BB1 = PnL)]
# bb2live <- BB2live[, .(date, BB2 = PnL)]
# tr1live <- TR1live[, .(date, TR1 = PnL)]
# 
# portfoliolive <- merge(bb1live, bb2live, by = "date", all = TRUE) 
# portfoliolive <- merge(portfoliolive, tr1live, by = "date", all = TRUE)
# portfoliolive[, number := as.integer(!is.na(BB1)) + as.integer(!is.na(BB2)) + as.integer(!is.na(TR1))]
# portfoliolive <- portfoliolive[number > 0]
# 
# portfoliolive[, BB1weight := ifelse(number < 3, 1/number, 0.3)]
# 
# portfoliolive[, BB2weight := ifelse(number < 3, 1/number, 0.5)]
# portfoliolive[, BB2weight := ifelse(number == 1, 0, BB2weight)]
# 
# portfoliolive[, TR1weight := ifelse(number < 3, 0, 0.2)]
# 
# portfoliolive[, BB1 := ifelse(is.na(BB1), 0, BB1)]
# portfoliolive[, BB2 := ifelse(is.na(BB2), 0, BB2)]
# portfoliolive[, TR1 := ifelse(is.na(TR1), 0, TR1)]
# 
# portfoliolive[, PnL := BB1 * BB1weight + BB2 * BB2weight + TR1 * TR1weight]


portfoliolive <- as.data.table(read_csv("C:/Users/mauritz/Box Sync/Polar Star - Company Folder/1. Polar Star/Analysis/Analysis and Research/code/Quant Portfolio/portfolioReturnsQCF.csv"))
portfoliolive <- portfoliolive[date >= "2018-06-04",.(date, PnL = return)]
portfoliolive[, type := ifelse(date < "2019-01-02", "Live", "PSQCF")]


```



```{r}

plotdata <- rbind(portfolio[date < portfoliolive[, min(date)],.(date, return = totalreturn, type = "Backtest")], portfoliolive[,.(date, return = PnL, type)])
plotdata[!is.na(return), cumret := cumsum(return)]

#write.csv(plotdata, "PSQCFreturns.csv", row.names = FALSE)

pic <- ggplot(plotdata, aes(x=date, y=cumret, group = type,
                            text = paste('Cummulative Return: ', round(cumret,2),
                                         '<br>Date: ', as.Date(date))
                            )
              ) +
  geom_line(aes(color = type)) +
  ylab("Cummulative Return") +
  xlab("") +
  ggtitle("Polar Star Quantitative Commodity Fund Equity Curve") +
  scale_x_date(breaks = pretty_breaks(n = 10)) +
  scale_y_continuous(breaks = pretty_breaks(n = 10)) +
  scale_color_manual(values = c(rgb(255, 90, 52, maxColorValue = 255), rgb(5,61,96, maxColorValue = 255), "firebrick")) +
  theme(
    plot.background = element_rect(fill = rgb(242,242,242, maxColorValue = 255)),
    panel.background = element_rect(fill = "white"),
    panel.grid.major.y = element_line(size = 0.5, linetype = 'solid',colour = "gray"),
    panel.grid.major.x = element_line(size = 0.5, linetype = 'dashed',colour = "gray"),
    axis.title.x = element_text(colour = rgb(5,61,96, maxColorValue = 255), size = 16),
    axis.title.y = element_text(colour = rgb(5,61,96, maxColorValue = 255), size = 16),
    plot.title = element_text(colour = rgb(5,61,96, maxColorValue = 255), size = 18, face = "bold"),
    axis.text = element_text(size=16),
    text=element_text(family="Century Gothic"),
    legend.key = element_rect(fill = rgb(242,242,242, maxColorValue = 255)),
    legend.background=element_blank(),
    legend.title=element_blank(),
    legend.position="bottom"
  )


ggplotly(pic, tooltip = c("text")) %>%
  layout(legend = list(orientation = "h", x = 0.4, y = -0.2))

```


## Quantitative Portfolio - Statistics {data-background="img/BlueBackgroundWithOrange.png"}


```{r}

portfolioReturns <- as.data.table(read_csv("C:/Users/mauritz/Box Sync/Polar Star - Company Folder/1. Polar Star/Analysis/Analysis and Research/code/Quant Portfolio/portfolioReturnsQCF.csv"))

portfolioReturns[, SnPnormalised := cumprod(SnPreturn + 1) - 1]

portfolioReturns[, SnPnormalised := SnPnormalised * 100]

##... FEES ...##

portfolioReturns[, year := year(date)]
portfolioReturns[, month := month(date)]
portfolioReturns[, half := ifelse(month < 7, 1, 2)]

portfolioReturns[, md := max(date), by = .(year, month)]

portfolioReturns[, management := 0]
portfolioReturns[date == md, management := 0.02/12]

portfolioReturns[, totalreturn := return - management]

portfolioReturns[, normalized := cumsum(totalreturn) * 100]


# s.data <- split(portfolioReturns, list(portfolioReturns$year, portfolioReturns$half))
# tflist <- map_lgl(s.data, function(df){ifelse(nrow(df) < 1, F, T)})
# s.data <- s.data[tflist]
# 
# 
# performancefees <- rbindlist(lapply(s.data, getPerformance))
# 
# portfolio <- merge(portfolio, performancefees, by = "date", all = TRUE)
# portfolio[, performance := ifelse(is.na(performance), 0, performance)]
# 
# portfolio[, totalreturn := totalreturn - performance]



##...      ...##

portfolioReturns[, flag := ifelse(date < "2019-01-02", "Live", "PSQCF")]
portfolioReturns <- portfolioReturns[, .(date, PSQCF = normalized, `S&P` = SnPnormalised, flag)]
portfolioReturns <- melt(portfolioReturns, id.vars=c("date", "flag"))
portfolioReturns[, flag := ifelse(variable == "S&P", "MSCI World TR Index", flag)]

#rebaldates <- as.Date(c("2018-07-03", "2018-08-02", "2018-09-05", "2018-10-02", "2018-11-02", "2018-12-02"))

pic <- ggplot(portfolioReturns, aes(x=date, y=value)) +
  geom_line(aes(color = flag), size = 1.2) +
  theme_bw() +
  scale_x_date(breaks = pretty_breaks(10)) +
  scale_y_continuous(breaks = pretty_breaks(10)) +
  ylab("Cummulative Return (%)") +
  #geom_text(data = portfolioReturns[date == max(date) | date %in% rebaldates], aes(label = round(normalized, 2)), nudge_y = 0.05) +
  ggtitle("Polar Star Quantitative Commodity Portfolio") +
  geom_hline(yintercept = 0, linetype = "dashed") +
  #geom_vline(xintercept = as.Date(rebaldates), linetype = "dashed") +
  #annotate(geom = "text", label = "Rebalance", x=as.Date(rebaldates), y=1.6, angle= 90)  +
  scale_color_manual(values = c(rgb(255, 90, 52, maxColorValue = 255), rgb(5,61,96, maxColorValue = 255), "darkgreen")) +
  #scale_linetype_manual(values = c("solid", "dashed", "solid")) +
  theme(
    plot.background = element_rect(fill = rgb(242,242,242, maxColorValue = 255)),
    panel.background = element_rect(fill = "white"),
    panel.grid.major.y = element_line(size = 0.5, linetype = 'solid',colour = "gray"),
    panel.grid.major.x = element_line(size = 0.5, linetype = 'dashed',colour = "gray"),
    axis.title.x = element_text(colour = rgb(5,61,96, maxColorValue = 255), size = 16),
    axis.title.y = element_text(colour = rgb(5,61,96, maxColorValue = 255), size = 16),
    plot.title = element_text(colour = rgb(5,61,96, maxColorValue = 255), size = 18, face = "bold"),
    axis.text = element_text(size=16),
    text=element_text(family="Century Gothic"),
    legend.key = element_rect(fill = rgb(242,242,242, maxColorValue = 255)),
    legend.background=element_blank(),
    legend.title=element_blank(),
    legend.position="bottom"
  )


ggplotly(pic)

```




## Quantitative Portfolio - Statistics {data-background="img/BlueBackgroundWithOrange.png"}

```{r}

meanrollingsharpe <- portfolio[, mean(`rolling return`, na.rm = TRUE)]


ggplot(portfolio, aes(x=`rolling return`)) +
  geom_histogram(aes(y=..density..), fill = rgb(255, 90, 52, maxColorValue = 255)) +
  ylab("Probability Density") +
  xlab("Rolling Return (%)") +
  ggtitle("Histogram of Polar Star Quantitative Commodity Fund 12 month rolling return") +
  scale_x_continuous(breaks = pretty_breaks(n = 10)) +
  scale_y_continuous(breaks = pretty_breaks(n = 10)) +
  geom_vline(xintercept = meanrollingsharpe, color = rgb(5,61,96, maxColorValue = 255), linetype = "dashed", size = 1.5) +
  theme(
    plot.background = element_rect(fill = rgb(242,242,242, maxColorValue = 255)),
    panel.background = element_rect(fill = "white"),
    panel.grid.major.y = element_line(size = 0.5, linetype = 'solid',colour = "gray"),
    panel.grid.major.x = element_line(size = 0.5, linetype = 'dashed',colour = "gray"),
    axis.title.x = element_text(colour = rgb(5,61,96, maxColorValue = 255), size = 16),
    axis.title.y = element_text(colour = rgb(5,61,96, maxColorValue = 255), size = 16),
    plot.title = element_text(colour = rgb(5,61,96, maxColorValue = 255), size = 13, face = "bold"),
    axis.text = element_text(size=16),
    text=element_text(family="Century Gothic")
  )

```


## Quantitative Portfolio - Statistics {data-background="img/BlueBackgroundWithOrange.png"}

```{r}

meanrollingsharpe <- portfolio[, mean(`rolling Sharpe`, na.rm = TRUE)]

ggplot(portfolio, aes(x=`rolling Sharpe`)) +
  geom_histogram(aes(y=..density..), fill = rgb(255, 90, 52, maxColorValue = 255)) +
  ylab("Probability Density") +
  xlab("Rolling Sharpe Ratio") +
  ggtitle("Histogram of Polar Star Quantitative Commodity Fund 12 month rolling Sharpe Ratio") +
  scale_x_continuous(breaks = pretty_breaks(n = 10)) +
  scale_y_continuous(breaks = pretty_breaks(n = 10)) +
  geom_vline(xintercept = meanrollingsharpe, color = rgb(5,61,96, maxColorValue = 255), linetype = "dashed", size = 1.5) +
  theme(
    plot.background = element_rect(fill = rgb(242,242,242, maxColorValue = 255)),
    panel.background = element_rect(fill = "white"),
    panel.grid.major.y = element_line(size = 0.5, linetype = 'solid',colour = "gray"),
    panel.grid.major.x = element_line(size = 0.5, linetype = 'dashed',colour = "gray"),
    axis.title.x = element_text(colour = rgb(5,61,96, maxColorValue = 255), size = 16),
    axis.title.y = element_text(colour = rgb(5,61,96, maxColorValue = 255), size = 16),
    plot.title = element_text(colour = rgb(5,61,96, maxColorValue = 255), size = 12, face = "bold"),
    axis.text = element_text(size=16),
    text=element_text(family="Century Gothic")
  )


```


## Quantitative Portfolio - Statistics {data-background="img/BlueBackgroundWithOrange.png"}

```{r}

tbldata <- rbind(portfolio[date < "2018-06-04",.(date, return = totalreturn)], portfoliolive[, .(date, return = PnL)])
tbldata[, month := month(date)]
tbldata[, year := year(date)]
tbldata[, cumret := cumprod(return + 1)]
tbldata[, cumret := cumret/tbldata[1, cumret]]
tbldata[, md := min(date), by = .(year, month)]

#monthlydata <- CalculateReturns(tbldata[date == md,.(date, cumret)])
#monthlydata <- as.data.table(monthlydata)
monthlydata <- tbldata[date != c("2018-06-01", "2018-06-02", "2018-06-03"), .(cumret = prod(return + 1) -1), by = .(year, month)]
monthlydata[, sign := sign(cumret)]

monthlydata[, index := paste(year, month, 1, sep = "-")]
monthlydata[, index := as.Date(index)]

avgneg <- monthlydata[sign == -1, mean(cumret) * 100]
avgpos <- monthlydata[sign == 1, mean(cumret) * 100]

noneg <- nrow(monthlydata[sign == -1])
nopos <- nrow(monthlydata[sign == 1])


tbl1 <- table.AnnualizedReturns(tbldata[, .(date, `1998-` = return)])
nms <- row.names(tbl1)
tbl1 <- as.data.table(tbl1)
tbl1[, variable := nms]

tbl1[variable != "Annualized Sharpe (Rf=0%)", `1998-` := `1998-` * 100]

maxdraw <- maxDrawdown(as.xts(tbldata[, .(date, return)]))[1] * 100
ret.maxdraw <- maxdraw/tbl1[variable == "Annualized Return", `1998-`]

parttoadd <- data.table(
  variable = c("Maximum Drawdown", "Maximum Drawdown/Annualized Return", "Number of Positive Months", "Number of Negative Months", "Average Positive Month Return", "Average Negative Month Return"),
  `1998-` = c(maxdraw, ret.maxdraw, nopos, noneg, avgpos, avgneg)
)

tbl1 <- rbind(tbl1, parttoadd)
tbl1 <- tbl1[,.(variable, `1998-`)]
tbl1[, `1998-` := round(`1998-`, 3)]

setnames(tbl1, "variable", "Statistic")


##

tbl2 <- table.AnnualizedReturns(tbldata[date >= "2008-01-01", .(date, `2008-` = return)])
nms <- row.names(tbl2)
tbl2 <- as.data.table(tbl2)
tbl2[, variable := nms]

tbl2[variable != "Annualized Sharpe (Rf=0%)", `2008-` := `2008-` * 100]

maxdraw <- maxDrawdown(as.xts(tbldata[date >= "2008-01-01", .(date, return)]))[1] * 100
ret.maxdraw <- maxdraw/tbl2[variable == "Annualized Return", `2008-`]

avgneg <- monthlydata[sign == -1 & index >= "2008-01-01", mean(cumret) * 100]
avgpos <- monthlydata[sign == 1 & index >= "2008-01-01", mean(cumret) * 100]

noneg <- nrow(monthlydata[sign == -1 & index >= "2008-01-01"])
nopos <- nrow(monthlydata[sign == 1 & index >= "2008-01-01"])


parttoadd <- data.table(
  variable = c("Maximum Drawdown", "Maximum Drawdown/Annualized Return", "Number of Positive Months", "Number of Negative Months", "Average Positive Month Return", "Average Negative Month Return"),
  `2008-` = c(maxdraw, ret.maxdraw, nopos, noneg, avgpos, avgneg)
)


tbl2 <- rbind(tbl2, parttoadd)
tbl2 <- tbl2[,.(variable, `2008-`)]
tbl2[, `2008-` := round(`2008-`, 3)]

setnames(tbl2, "variable", "Statistic")


##


tbl3 <- table.AnnualizedReturns(tbldata[date >= "2015-01-01", .(date, `2015-` = return)])
nms <- row.names(tbl3)
tbl3 <- as.data.table(tbl3)
tbl3[, variable := nms]

tbl3[variable != "Annualized Sharpe (Rf=0%)", `2015-` := `2015-` * 100]

maxdraw <- maxDrawdown(as.xts(tbldata[date >= "2015-01-01", .(date, return)]))[1] * 100
ret.maxdraw <- maxdraw/tbl3[variable == "Annualized Return", `2015-`]

avgneg <- monthlydata[sign == -1 & index >= "2015-01-01", mean(cumret) * 100]
avgpos <- monthlydata[sign == 1 & index >= "2015-01-01", mean(cumret) * 100]

noneg <- nrow(monthlydata[sign == -1 & index >= "2015-01-01"])
nopos <- nrow(monthlydata[sign == 1 & index >= "2015-01-01"])


parttoadd <- data.table(
  variable = c("Maximum Drawdown", "Maximum Drawdown/Annualized Return", "Number of Positive Months", "Number of Negative Months", "Average Positive Month Return", "Average Negative Month Return"),
  `2015-` = c(maxdraw, ret.maxdraw, nopos, noneg, avgpos, avgneg)
)


tbl3 <- rbind(tbl3, parttoadd)
tbl3 <- tbl3[,.(variable, `2015-`)]
tbl3[, `2015-` := round(`2015-`, 3)]

setnames(tbl3, "variable", "Statistic")


##

tbl4 <- table.AnnualizedReturns(tbldata[date >= "2018-06-04", .(date, live = return)])
nms <- row.names(tbl4)
tbl4 <- as.data.table(tbl4)
tbl4[, variable := nms]

tbl4[variable != "Annualized Sharpe (Rf=0%)", live := live * 100]

maxdraw <- maxDrawdown(as.xts(tbldata[date >= "2018-06-04", .(date, return)]))[1] * 100
ret.maxdraw <- maxdraw/tbl4[variable == "Annualized Return", live]

avgneg <- monthlydata[sign == -1 & index >= "2018-06-04", mean(cumret) * 100]
avgpos <- monthlydata[sign == 1 & index >= "2018-06-04", mean(cumret) * 100]

noneg <- nrow(monthlydata[sign == -1 & index >= "2018-06-04"])
nopos <- nrow(monthlydata[sign == 1 & index >= "2018-06-04"])


parttoadd <- data.table(
  variable = c("Maximum Drawdown", "Maximum Drawdown/Annualized Return", "Number of Positive Months", "Number of Negative Months", "Average Positive Month Return", "Average Negative Month Return"),
  live = c(maxdraw, ret.maxdraw, nopos, noneg, avgpos, avgneg)
)


tbl4 <- rbind(tbl4, parttoadd)
tbl4 <- tbl4[,.(variable, live)]
tbl4[, live := round(live, 3)]

setnames(tbl4, "variable", "Statistic")


##

tbl <- merge(tbl1, tbl2, by = "Statistic")
tbl <- merge(tbl, tbl3, by = "Statistic")
tbl <- merge(tbl, tbl4, by = "Statistic")

formattable(tbl)

```


# Quantitative Portfolio as supplement to S&P500 {data-background="img/background_mtns_1.png"}

## {data-background="img/BlueBackgroundWithOrange.png"}

```{r}

snp <- as.data.table(read_csv("C:/Users/mauritz/Box Sync/Polar Star - Company Folder/1. Polar Star/Analysis/Analysis and Research/code/Quant Portfolio/spx.csv"))

snp <- snp[date >= portfolio[, min(date)]]
snp[, SnP := SnP/snp[1, SnP]]
setnames(snp, "SnP", "S&P500")

quant <- portfolio[, .(date, PSQCF = normalised)]

together <- merge(quant, snp, by = "date")

```




```{r}

together[, month := month(date)]
together[, year := year(date)]
together[, md := max(date), by = .(year, month)]
together <- together[date == md]
together <- together[, .(date, PSQCF, `S&P500`)]

returns <- CalculateReturns(together)

portion <- seq(0, 1, 0.01)

part <- vector(mode = "list", length = length(portion))
for(i in 1:length(portion)){
  #i=1
  tempret <- Return.portfolio(returns, weights = c(portion[i], 1-portion[i]))
  tempret <- as.data.table(tempret)
  tempret[, PS_ratio := portion[i]]
  part[[i]] <- tempret  
}

allparts <- rbindlist(part)
allparts[, label := paste("Polar Star", PS_ratio * 100)]

widedata <- dcast(allparts, index ~ label, value.var = "portfolio.returns")
widedata <- as.xts(widedata)


cols <- paste("Polar Star", seq(0, 100, 10))

draws <- maxDrawdown(widedata)
stdevs <- StdDev(widedata)

annData <- table.AnnualizedReturns(widedata, scale = 12, digits = 8)
annData <- as.data.table(annData[1,])
annData <- rbind(annData, stdevs, draws)
annData[, variable := c("Annualised Return", "Standard Deviation", "Worst Drawdown")]


plotdata <- melt(annData, id.vars = "variable", variable.name = "PSFraction")
plotdata <- dcast(plotdata, PSFraction ~ variable, value.var = "value")
setnames(plotdata, names(plotdata), gsub(" ", "_", names(plotdata)))
plotdata[, Annualised_Return := Annualised_Return * 100]
plotdata[, Standard_Deviation := Standard_Deviation * 100]
plotdata[, Worst_Drawdown := Worst_Drawdown * 100]
plotdata[, marker := tstrsplit(PSFraction ,"Star", keep = 2)]
plotdata[, marker := as.integer(marker)]
plotdata <- plotdata[order(marker)]
plotdata[, group := ifelse(marker %% 10 == 0, "A", "B")]
plotdata[, PSFraction := paste0(PSFraction, "%")]

plotdata[, PSFraction := gsub("Polar Star", "PSQCF", PSFraction)]

```



```{r}

ggplot(plotdata, aes(x=Standard_Deviation, y=Annualised_Return)) +
  geom_point(aes(color = group, size = group)) +
  geom_text(data = plotdata[group == "A"], 
            aes(label = PSFraction), nudge_x = 0.15, size = 4.5, colour = rgb(5,61,96, maxColorValue = 255) ) +
  ylab("Annualised Return") +
  xlab("Standard Deviation") +
  ggtitle("Polar Star Quantitiave Commodity Fund and S&P500 Efficeint Frontier") +
  scale_size_manual(values = c(5, 1), guide=FALSE) +
  scale_color_manual(values = c(rgb(255, 90, 52, maxColorValue = 255), "gray"), guide=FALSE) +
  scale_x_continuous(breaks = pretty_breaks(n = 10), expand = c(.1, 0)) +
  scale_y_continuous(breaks = pretty_breaks(n = 10)) +
  theme(
    plot.background = element_rect(fill = rgb(242,242,242, maxColorValue = 255)),
    panel.background = element_rect(fill = "white"),
    panel.grid.major.y = element_line(size = 0.5, linetype = 'solid',colour = "gray"),
    panel.grid.major.x = element_line(size = 0.5, linetype = 'dashed',colour = "gray"),
    axis.title.x = element_text(colour = rgb(5,61,96, maxColorValue = 255), size = 16),
    axis.title.y = element_text(colour = rgb(5,61,96, maxColorValue = 255), size = 16),
    plot.title = element_text(colour = rgb(5,61,96, maxColorValue = 255), size = 12, face = "bold"),
    axis.text = element_text(size=16),
    text=element_text(family="Century Gothic")
  )



```


## Portfolio Comparison {data-background="img/BlueBackgroundWithOrange.png"}

```{r}

allPortfolios <- together[, .(date, `S&P500`, PSQCF, `PSQCF and S&P500` = 0.3 * PSQCF + 0.7 * `S&P500`)]


monthlydata <- CalculateReturns(allPortfolios)
monthlydata <- as.data.table(monthlydata)

monthlydata <- melt(monthlydata, id.vars = "index", variable.name = "Portfolio", value.name = "Return")
monthlydata <- monthlydata[!is.na(Return)]
monthlydata[, sign := sign(Return)]

avgrets <- monthlydata[, .(avgret = mean(Return) * 100, .N), by = .(Portfolio, sign)]

numbers <- dcast(avgrets, sign ~ Portfolio, value.var = "N")
numbers[, variable := ifelse(sign == -1, "Number of Negative Months", "Number of Positive Months")]
numbers[, sign := NULL]

posnegrets <- dcast(avgrets, sign ~ Portfolio, value.var = "avgret")
posnegrets[, variable := ifelse(sign == -1, "Avereage Negative Month Return", "Average Positive Month Return")]
posnegrets[, sign := NULL]

tbl1 <- table.AnnualizedReturns(as.xts(CalculateReturns(allPortfolios)))
maxdraw <- maxDrawdown(as.xts(CalculateReturns(allPortfolios))) 


tbl1 <- rbind(tbl1, maxdraw)
nms <- row.names(tbl1)
tbl1 <- as.data.table(tbl1)
tbl1[, variable := nms]

tbl1[variable != "Annualized Sharpe (Rf=0%)", `PSQCF and S&P500` := `PSQCF and S&P500` * 100]
tbl1[variable != "Annualized Sharpe (Rf=0%)", `S&P500` := `S&P500` * 100]
tbl1[variable != "Annualized Sharpe (Rf=0%)", PSQCF := PSQCF * 100]


tbl <- rbind(tbl1, numbers, posnegrets)
setnames(tbl, "variable", "Statistic")

tempdata <- melt(tbl, id.vars = "Statistic", variable.name = "Portfolio")
tempdata[, value := round(value, 3)]

tbl <- dcast(tempdata, Statistic ~ Portfolio, value.var = "value")
tbl <- tbl[, .(Statistic, `S&P500`, PSQCF, `PSQCF and S&P500`)]


formattable(tbl)

```



# Polar Star Multi Strategy Portfolio {data-background="img/background_mtns_1.png"}

## Polar Star Products and S&P500 {data-background="img/BlueBackgroundWithOrange.png"}

```{r}

together <- as.data.table(read_csv("C:/Users/mauritz/Box Sync/Polar Star - Company Folder/1. Polar Star/Analysis/Analysis and Research/code/Quant Portfolio/data/together.csv"))
together <- together[order(date)]

fees <- portfolio[date == md, .(year, month, fees = management + performance)]

together[, year := year(date)]
together[, month := month(date)]
together <- merge(together, fees, by = c("year", "month"))
together[, PSQCF := PSQCF - fees]
together[ , c("year", "month", "fees") := NULL ]


together[, PSLtd := cumprod(1+PSLtd)]
together[, PSQCF := cumprod(1+PSQCF)]
together[, SnP := cumprod(1+SnP)]

together[, PSLtd := PSLtd/together[1, PSLtd]]
together[, PSQCF := PSQCF/together[1, PSQCF]]
together[, SnP := SnP/together[1, SnP]]

plotdata <- melt(together, id.vars = "date")
plotdata <- plotdata[order(date)]


pic <- ggplot(plotdata, aes(x=date, y=value, group = variable,
                     text = paste('Cummulative Return: ', round(value,2),
                                  '<br>Date: ', as.Date(date))
                     )
       ) +
  geom_line(aes(group = variable, color = variable), size = 1.2) +
  ylab("Cummulative Return") +
  xlab("") +
  ggtitle("Polar Star Quantitative Commodity Fund, Ltd an S&P500 Equity Curves") +
  scale_x_date(breaks = pretty_breaks(n = 10)) +
  scale_y_continuous(breaks = pretty_breaks(n = 10)) +
  theme(
    plot.background = element_rect(fill = rgb(242,242,242, maxColorValue = 255)),
    panel.background = element_rect(fill = "white"),
    panel.grid.major.y = element_line(size = 0.5, linetype = 'solid',colour = "gray"),
    panel.grid.major.x = element_line(size = 0.5, linetype = 'dashed',colour = "gray"),
    axis.title.x = element_text(colour = rgb(5,61,96, maxColorValue = 255), size = 16),
    axis.title.y = element_text(colour = rgb(5,61,96, maxColorValue = 255), size = 16),
    plot.title = element_text(colour = rgb(5,61,96, maxColorValue = 255), size = 12, face = "bold"),
    axis.text = element_text(size=16),
    text=element_text(family="Century Gothic"),
    legend.key = element_rect(fill = rgb(242,242,242, maxColorValue = 255)),
    legend.background=element_blank(),
    legend.title=element_blank(),
    legend.position="bottom"
  )


ggplotly(pic) %>%
  layout(legend = list(orientation = "h", x = 0.4, y = -0.2))

```



## Combine Discretionary and Systematic Portfolios {data-background="img/BlueBackgroundWithOrange.png"}

```{r}

together <- as.data.table(read_csv("C:/Users/mauritz/Box Sync/Polar Star - Company Folder/1. Polar Star/Analysis/Analysis and Research/code/Quant Portfolio/data/together.csv"))

fees <- portfolio[date == md, .(year, month, fees = management + performance)]

together[, year := year(date)]
together[, month := month(date)]
together <- merge(together, fees, by = c("year", "month"))
together[, PSQCF := PSQCF - fees]
together[ , c("year", "month", "fees") := NULL ]


returns <- as.xts(together[,.(date, PSLtd, PSQCF)])

portion <- seq(0, 1, 0.01)

part <- vector(mode = "list", length = length(portion))
for(i in 1:length(portion)){
  #i=1
  tempret <- Return.portfolio(returns, weights = c(portion[i], 1-portion[i]))
  tempret <- as.data.table(tempret)
  tempret[, PS_ratio := portion[i]]
  part[[i]] <- tempret  
}

allparts <- rbindlist(part)
allparts[, label := paste("Polar Star", PS_ratio * 100)]

widedata <- dcast(allparts, index ~ label, value.var = "portfolio.returns")
widedata <- as.xts(widedata)


cols <- paste("Polar Star", seq(0, 100, 10))

draws <- maxDrawdown(widedata)
stdevs <- StdDev(widedata)

annData <- table.AnnualizedReturns(widedata, scale = 12, digits = 8)
annData <- as.data.table(annData[1,])
annData <- rbind(annData, stdevs, draws)
annData[, variable := c("Annualised Return", "Standard Deviation", "Worst Drawdown")]


plotdata <- melt(annData, id.vars = "variable", variable.name = "PSFraction")
plotdata <- dcast(plotdata, PSFraction ~ variable, value.var = "value")
setnames(plotdata, names(plotdata), gsub(" ", "_", names(plotdata)))
plotdata[, Annualised_Return := Annualised_Return * 100]
plotdata[, Standard_Deviation := Standard_Deviation * 100]
plotdata[, Worst_Drawdown := Worst_Drawdown * 100]
plotdata[, marker := tstrsplit(PSFraction ,"Star", keep = 2)]
plotdata[, marker := as.integer(marker)]
plotdata <- plotdata[order(marker)]
plotdata[, group := ifelse(marker %% 10 == 0, "A", "B")]
plotdata[, PSFraction := paste0(PSFraction, "%")]

plotdata[, PSFraction := gsub("Polar Star", "PS Ltd", PSFraction)]

```


```{r}

ggplot(plotdata, aes(x=Standard_Deviation, y=Annualised_Return)) +
  geom_point(aes(color = group, size = group)) +
  geom_text(data = plotdata[group == "A"], 
            aes(label = PSFraction), nudge_x = 0.15, size = 4.5, colour = rgb(5,61,96, maxColorValue = 255) ) +
  ylab("Annualised Return") +
  xlab("Standard Deviation") +
  ggtitle("Polar Star Ltd and Polar Star Quantitiave Commodity Fund Efficeint Frontier") +
  scale_size_manual(values = c(5, 1), guide=FALSE) +
  scale_color_manual(values = c(rgb(255, 90, 52, maxColorValue = 255), "gray"), guide=FALSE) +
  scale_x_continuous(breaks = pretty_breaks(n = 10), expand = c(.1, 0)) +
  scale_y_continuous(breaks = pretty_breaks(n = 10)) +
  theme(
    plot.background = element_rect(fill = rgb(242,242,242, maxColorValue = 255)),
    panel.background = element_rect(fill = "white"),
    panel.grid.major.y = element_line(size = 0.5, linetype = 'solid',colour = "gray"),
    panel.grid.major.x = element_line(size = 0.5, linetype = 'dashed',colour = "gray"),
    axis.title.x = element_text(colour = rgb(5,61,96, maxColorValue = 255), size = 16),
    axis.title.y = element_text(colour = rgb(5,61,96, maxColorValue = 255), size = 16),
    plot.title = element_text(colour = rgb(5,61,96, maxColorValue = 255), size = 12, face = "bold"),
    axis.text = element_text(size=16),
    text=element_text(family="Century Gothic")
  )



```





## Polar Star Multi Strategy as enhancement to S&P500 {data-background="img/BlueBackgroundWithOrange.png"}

```{r}

allPortfolios <- together[, .(date, `50-50` = PSLtd * 0.5 + PSQCF * 0.5, SnP, PSLtd, PSQCF)]


returns <- as.xts(allPortfolios[,.(date, `50-50`, SnP)])

portion <- seq(0, 1, 0.01)

part <- vector(mode = "list", length = length(portion))
for(i in 1:length(portion)){
  #i=1
  tempret <- Return.portfolio(returns, weights = c(portion[i], 1-portion[i]))
  tempret <- as.data.table(tempret)
  tempret[, PS_ratio := portion[i]]
  part[[i]] <- tempret  
}

allparts <- rbindlist(part)
allparts[, label := paste("Polar Star", PS_ratio * 100)]

widedata <- dcast(allparts, index ~ label, value.var = "portfolio.returns")
widedata <- as.xts(widedata)


cols <- paste("Polar Star", seq(0, 100, 10))

draws <- maxDrawdown(widedata)
stdevs <- StdDev(widedata)

annData <- table.AnnualizedReturns(widedata, scale = 12, digits = 8)
annData <- as.data.table(annData[1,])
annData <- rbind(annData, stdevs, draws)
annData[, variable := c("Annualised Return", "Standard Deviation", "Worst Drawdown")]


plotdata <- melt(annData, id.vars = "variable", variable.name = "PSFraction")
plotdata <- dcast(plotdata, PSFraction ~ variable, value.var = "value")
setnames(plotdata, names(plotdata), gsub(" ", "_", names(plotdata)))
plotdata[, Annualised_Return := Annualised_Return * 100]
plotdata[, Standard_Deviation := Standard_Deviation * 100]
plotdata[, Worst_Drawdown := Worst_Drawdown * 100]
plotdata[, marker := tstrsplit(PSFraction ,"Star", keep = 2)]
plotdata[, marker := as.integer(marker)]
plotdata <- plotdata[order(marker)]
plotdata[, group := ifelse(marker %% 10 == 0, "A", "B")]
plotdata[, PSFraction := paste0(PSFraction, "%")]

plotdata[, PSFraction := gsub("Polar Star", "PS Multi Strat", PSFraction)]


```


```{r}

ggplot(plotdata, aes(x=Standard_Deviation, y=Annualised_Return)) +
  geom_point(aes(color = group, size = group)) +
  geom_text(data = plotdata[group == "A"], 
            aes(label = PSFraction), nudge_x = 0.25, size = 4.5, colour = rgb(5,61,96, maxColorValue = 255) ) +
  ylab("Annualised Return") +
  xlab("Standard Deviation") +
  ggtitle("Polar Star Multi Strategy and S&P500 Efficeint Frontier") +
  scale_size_manual(values = c(5, 1), guide=FALSE) +
  scale_color_manual(values = c(rgb(255, 90, 52, maxColorValue = 255), "gray"), guide=FALSE) +
  scale_x_continuous(breaks = pretty_breaks(n = 10), expand = c(.1, 0)) +
  scale_y_continuous(breaks = pretty_breaks(n = 10)) +
  theme(
    plot.background = element_rect(fill = rgb(242,242,242, maxColorValue = 255)),
    panel.background = element_rect(fill = "white"),
    panel.grid.major.y = element_line(size = 0.5, linetype = 'solid',colour = "gray"),
    panel.grid.major.x = element_line(size = 0.5, linetype = 'dashed',colour = "gray"),
    axis.title.x = element_text(colour = rgb(5,61,96, maxColorValue = 255), size = 16),
    axis.title.y = element_text(colour = rgb(5,61,96, maxColorValue = 255), size = 16),
    plot.title = element_text(colour = rgb(5,61,96, maxColorValue = 255), size = 12, face = "bold"),
    axis.text = element_text(size=16),
    text=element_text(family="Century Gothic")
  )



```




## Portfolio Comparison {data-background="img/BlueBackgroundWithOrange.png"}

```{r}

allPortfolios[, `PS Multi Strategy and S&P500` := `50-50` * 0.5 + SnP * 0.5]
setnames(allPortfolios, c("50-50", "SnP"), c("PS Multi Strategy","S&P500"))


monthlydata <- as.xts(allPortfolios)
monthlydata <- as.data.table(monthlydata)
monthlydata <- melt(monthlydata, id.vars = "index", variable.name = "Portfolio", value.name = "Return")
monthlydata[, sign := sign(Return)]

avgrets <- monthlydata[, .(avgret = mean(Return) * 100, .N), by = .(Portfolio, sign)]

numbers <- dcast(avgrets, sign ~ Portfolio, value.var = "N")
numbers[, variable := ifelse(sign == -1, "Number of Negative Months", "Number of Positive Months")]
numbers[, sign := NULL]

posnegrets <- dcast(avgrets, sign ~ Portfolio, value.var = "avgret")
posnegrets[, variable := ifelse(sign == -1, "Avereage Negative Month Return", "Average Positive Month Return")]
posnegrets[, sign := NULL]

tbl1 <- table.AnnualizedReturns(as.xts(allPortfolios))
maxdraw <- maxDrawdown(as.xts(allPortfolios))

tbl1 <- rbind(tbl1, maxdraw)
nms <- row.names(tbl1)
tbl1 <- as.data.table(tbl1)
tbl1[, variable := nms]

tbl1[variable != "Annualized Sharpe (Rf=0%)", PSLtd := PSLtd * 100]
tbl1[variable != "Annualized Sharpe (Rf=0%)", `S&P500` := `S&P500` * 100]
tbl1[variable != "Annualized Sharpe (Rf=0%)", PSQCF := PSQCF * 100]
tbl1[variable != "Annualized Sharpe (Rf=0%)", `PS Multi Strategy` := `PS Multi Strategy` * 100]
tbl1[variable != "Annualized Sharpe (Rf=0%)", `PS Multi Strategy and S&P500`:= `PS Multi Strategy and S&P500` * 100]


tbl <- rbind(tbl1, numbers, posnegrets)
setnames(tbl, "variable", "Statistic")

tempdata <- melt(tbl, id.vars = "Statistic", variable.name = "Portfolio")
tempdata[, value := round(value, 3)]

tbl <- dcast(tempdata, Statistic ~ Portfolio, value.var = "value")
tbl <- tbl[, .(Statistic, `S&P500`, `PS Multi Strategy`, `PS Multi Strategy and S&P500`)]


formattable(tbl)

```


# Summary {data-background="img/background_mtns_1.png"}


## {data-background="img/BlueBackgroundWithOrange.png"}


Combining a 

* discretionary and 
* systematic approach 
 
to investing in commodities we create a product with
 
* positive expected return which is
* uncorrelated to equities
 
that gives superior risk adjusted returns.

